### 2019-08-17 Analysis/GWAS_Prep.Rmd

---
title: "GWAS_Prep"
output:
  html_document:
    toc: true
    # toc_float: true
    # theme: united
    theme: cerulean
    # number_sections: true
date: "`r Sys.Date()`"
author: "Mikhail Dozmorov"
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
# Set up the environment
library(knitr)
opts_chunk$set(cache.path='cache/', fig.path='img/', cache=F, tidy=T, fig.keep='high', echo=F, dpi=100, warnings=F, message=F, comment=NA, warning=F, results='as.is', fig.width = 10, fig.height = 6) #out.width=700, 
library(pander)
panderOptions('table.split.table', Inf)
panderOptions('knitr.auto.asis', FALSE)
set.seed(1)
library(dplyr)
options(stringsAsFactors = FALSE)
```

```{r}
#Read in data and packages

library(gwascat)
library(limma)
library(dplyr)
library(DO.db)
library(DBI)
library(MDmisc)
library(ontoCAT)
library(rowr)
require(readr)

data(ebicat37)
```

```{r}
#Initial summary

#Convert to data frame

ebi_dat = as.data.frame(mcols(ebicat37))

#Put comma seperated genes on different lines

ebi_dat = unembed(ebi_dat, "MAPPED_GENE", sep = ",")
ebi_dat = unembed(ebi_dat, "MAPPED_GENE", sep = " - ")

#Remove extra whitespace from unmebed

ebi_dat$MAPPED_GENE = gsub(" ", "", ebi_dat$MAPPED_GENE)

#Summarise

init_sum = ebi_dat %>% dplyr::select(DISEASE.TRAIT, MAPPED_GENE, SNPS)  %>% distinct() %>% group_by(DISEASE.TRAIT) %>% summarise(snp_count_total = length(unique(SNPS)), snp_count_no_gene = sum(MAPPED_GENE == ""), snp_prop_nogene = snp_count_no_gene/snp_count_total,gene_count_total = length(unique(MAPPED_GENE)) - sum(unique(MAPPED_GENE =="")))

write.csv(init_sum, "supplementary_table_diseases_raw.csv", row.names = FALSE)

```

```{r }
#Converting aliases to gene names

alias2SymbolEbicat = function(ebicat37) {

  ebi_dat = as.data.frame(mcols(ebicat37))

  #Unembed comma and - seperated values

  ebi_dat = unembed(ebi_dat, "MAPPED_GENE", sep = ",")
  ebi_dat = unembed(ebi_dat, "MAPPED_GENE", sep = " - ")

  #Remove extra spaces

  ebi_dat$MAPPED_GENE = gsub(" ", "", ebi_dat$MAPPED_GENE)

  #Summarise the results

  #Keeping track of changes (sapply necessary to keep placeholders for unmapped genes)

  recomb_genes = sapply(ebi_dat$MAPPED_GENE, function(x) alias2Symbol(x, species = "Hs"))

  recomb_genes = unlist(lapply(recomb_genes, function(x) paste0(x,collapse=' ')))

  change = (ebi_dat$MAPPED_GENE == recomb_genes)

  sum_vec = ifelse(change == TRUE, "Same", ifelse(change == FALSE & recomb_genes == "", "No Match", "Converted"))

  ebi_dat$MAPPED_GENE = recomb_genes

  return(ebi_dat)
}

ebi_dat = alias2SymbolEbicat(ebicat37)
```

```{r}
#Pulling out useful info

main_inf = ebi_dat %>% dplyr::select(MAPPED_GENE, SNPS, CHR_ID, CHR_POS, INITIAL.SAMPLE.DESCRIPTION, MAPPED_TRAIT, MAPPED_TRAIT_URI, DISEASE.TRAIT)

#Updated EFO table

```

```{r}
#Using EFO IDs from ebicat to get a hierarchy

efo_loc = "https://raw.githubusercontent.com/EBISPOT/efo/master/efo.owl"

efo = getOntology(efo_loc)
# save(list = "efo",file =  "./data/efo.rdata")
# 
# load("./data/efo.rdata")

#Cleaning up efo data to remove urls and putting comma seperated values on new lines

main_inf = unembed(main_inf, "MAPPED_TRAIT_URI", sep = ",")
main_inf$MAPPED_TRAIT_URI = basename(main_inf$MAPPED_TRAIT_URI)

tree = c()
for (i in unique(main_inf$MAPPED_TRAIT_URI)) {

  #Get entire EFO tree

  parents = getAllTermParentsById(efo, i)

  #Pull out labels and make into vector

  parents = sapply(parents, getLabel)

  #Cbind and fill in non-matching entries with NA

  tree = rowr::cbind.fill(tree,parents, fill = NA)

}

#Remove original column

tree = tree[,-1]

colnames(tree) = unique(main_inf$MAPPED_TRAIT_URI)

#Pick out important disorder groups

imp_cats = c("Abnormality of refraction", "Abnormality of the eye", "age", "anthropometric measurement", "auditory system disease", "autoimmune disease", "body weights and measures", "bone measurement", "cancer", "carcinoma", "cardiovascular disease", "cardiovascular disease biomarker measurement", "cardiovascular measurement", "diabetes mellitus", "diabetes mellitus biomarker", "digestive system disease", "eye measurement", "glucose measurement", "head and neck disorder", "heart function measurement", "hematological measurement", "immune system disease", "infectious disease biomarker", "insulin measurement", "lipid measurement", "lipid or lipoprotein measurement", "lipoprotein measurement", "mental or behavioural disorder", "metabolic disease", "metabolite measurement", "neoplasm", "nervous system disease", "neurodegenerative disease", "pulmonary function measurement", "reproductive system disease", "respiratory disease biomarker", "skin disease", "sleep measurement", "temporal measurement")

#Output cats by pipe

#Get categories for each unique EFO ID

cat_ass = apply(tree, 2, function(x) c(na.omit(match(x, imp_cats))))

#Replace categoryless values with NA

cat_ass = lapply(cat_ass, function(x) {
  if (length(x)<1) {
    cat = NA 
  } else {
    cat = x
  }
  return(cat)
})

#Assign category names

mapped_groups = stack(cat_ass)
mapped_groups$values = imp_cats[mapped_groups$values]
colnames(mapped_groups) = c("Category", "MAPPED_TRAIT_URI")

#Add groups to original data frame

main_inf = left_join(main_inf,mapped_groups, by = "MAPPED_TRAIT_URI")

main_inf$Category = as.character(main_inf$Category)

main_inf = main_inf %>% group_by(MAPPED_TRAIT_URI) %>% mutate(Category = paste0(unique(Category), collapse = "|")) %>% distinct()

#Collapse genes and disease names for simplifcation purposes and then get unique combinations

names_genes = main_inf %>% ungroup() %>% dplyr::select(DISEASE.TRAIT, MAPPED_GENE) %>% distinct()

#Loop through disease traits and print a gene list for each

dir.create('./data/Disease_Genes')

dis_dir = "./data/Disease_Genes/"

for (i in unique(names_genes$DISEASE.TRAIT)) {
genes_curr = names_genes %>% filter(DISEASE.TRAIT == i & MAPPED_GENE != "") %>% dplyr::select(DISEASE.TRAIT, MAPPED_GENE) %>% mutate(DISEASE.TRAIT = gsub("\\/", "-", DISEASE.TRAIT), DISEASE.TRAIT = gsub("\\*", "-", DISEASE.TRAIT), File_Path =  DISEASE.TRAIT[1]) %>% ungroup() %>% dplyr::select(DISEASE.TRAIT, MAPPED_GENE, File_Path) %>% distinct()

genes_disease = genes_curr %>% ungroup() %>% dplyr::select(MAPPED_GENE) %>% distinct()

genes_curr = unique(genes_curr)

write.table(genes_disease, paste0(dis_dir, substr(genes_curr$File_Path[1],1,100), ".txt"), row.names = FALSE, col.names = FALSE)
}

#Create map of EFO ids to terms

EFO_list = lapply(unique(main_inf$MAPPED_TRAIT_URI), function(x) {
  if (x == "GO_190252") {
    term = NA
  } else {
  term = getLabel(getTermById(efo, x))
  }
  return(term)
  })

names(EFO_list) = unique(main_inf$MAPPED_TRAIT_URI)

EFO_Frame = stack(EFO_list)

colnames(EFO_Frame) = c("EFO_Name", "MAPPED_TRAIT_URI")

main_inf = left_join(main_inf, EFO_Frame, "MAPPED_TRAIT_URI")

```

```{r}
main_sum = main_inf %>% ungroup() %>%  dplyr::select(DISEASE.TRAIT, MAPPED_GENE, Category, SNPS)  %>% distinct() %>% group_by(DISEASE.TRAIT) %>% mutate(snp_count_total = length(unique(SNPS)), snp_count_no_gene = sum(MAPPED_GENE == ""), snp_prop_nogene = snp_count_no_gene/snp_count_total,gene_count_total = length(unique(MAPPED_GENE))) %>% dplyr::select(DISEASE.TRAIT, Category, snp_count_total, snp_count_no_gene, snp_prop_nogene, gene_count_total) %>% distinct()

write.csv(main_sum, "supplementary_table_diseases_eof.csv", row.names = FALSE)
```

```{r}
#Creating KEGG pathways

mart  = useMart("ensembl", dataset="hsapiens_gene_ensembl")
min_genes = 5

xx = as.list(KEGGPATHID2EXTID)     # All KEGG pathways
xx = xx[ grepl("hsa", names(xx)) ] # Keep hsa pathways only
xx_length = sapply(xx, length)     # Number of genes in each pathway
xx = xx[xx_length > min_genes]     # Keep pathways with minimum number of genes or higher
xx_sample = sample(1:length(xx), size = length(xx), replace = FALSE) # reshuffled indices
# Get all gene symbols for KEGG pathways
genes = getBM(attributes=c('entrezgene', 'hgnc_symbol'), filters='entrezgene', values=unique(unlist(xx)), mart=mart, uniqueRows=T)

kegg_paths = lapply(xx, function(x) left_join(data.frame(entrezgene = as.integer(x)), genes))

#Loop through kegg pathways and save genes

dir.create("./data/Disease_Genes/KEGG")

for (i in 1:length(kegg_paths)) {
  genes = kegg_paths[[i]][,2]
  write.table(genes, paste0("./data/Disease_Genes/KEGG/", names(kegg_paths)[i] ), row.names = FALSE, col.names = FALSE)

}

```