### 2019-08-17 manuscript/Analysis_Tables_New.Rmd

---
title: "Secondary analysis of tables"
# author: "Author's Name"
# date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: no
  html_document:
    theme: cerulean
    toc: yes
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
# Set up the environment
library(knitr)
opts_chunk$set(cache.path='cache/', fig.path='img/', cache=F, tidy=T, fig.keep='high', echo=F, dpi=100, warnings=F, message=F, comment=NA, warning=F, results='as.is', fig.width = 10, fig.height = 6) #out.width=700, 
library(pander)
panderOptions('table.split.table', Inf)
set.seed(1)
library(dplyr)
options(stringsAsFactors = FALSE)
```

```{r}
library(readr)
library(ggplot2)
library(ggrepel)
library(gridExtra)
source("https://raw.githubusercontent.com/mdozmorov/MDmisc/master/R/venn.R")
library(corrplot)
col3 <- colorRampPalette(c("blue", "white", "red"))
library(openxlsx)
library(cowplot)
```

```{r}
max_genes <- 200 # Maximum number of genes for a network to be included

fileNameIn0  <- "tables/supplementary_table_S1.csv"   # diseases and traits, coherence new KEGG, REACTOME, CC. 2019-07-14, Kellen
# iconv -f UTF-8 -t ISO-8859-1 -c supplementary_table_S1.csv > tmp.csv
mtx0 <- read_csv(fileNameIn0) %>% as.data.frame()
rownames(mtx0) <- mtx0$Disease
```

### Coherence distribution

```{r}
mtx0[, "String Filt Coherence KEGG"] %>% unlist %>% c %>% unname %>% hist
mtx0[, "String Filt Coherence KEGG"] %>% unlist %>% c %>% unname %>% summary
```

### FigS1_corr_coherence.png

**Supplementary Figure S1.[FILE: Figures/FigS1_corr_coherence.png] Normalized coherence is unaffected by network size.** Red/blue gradient and numbers represent Pearson correlation coefficients between corresponding entries.

```{r}
# Selected data for correlation
# "Disease", "Category", "Number of SNPs", "Number of genes", "Biogrid Coherence", "String Coherence", "String Filt Coherence", "Biogrid Coherence Reactome", "String Coherence Reactome", "String Filt Coherence Reactome""Biogrid Coherence CC", "String Coherence CC", "String Filt Coherence CC", "Biogrid P-value", "String P-value", "String Filt P-value" 
mtx0_cor <- mtx0 %>% dplyr::select(c("Number of SNPs", "Number of genes", "Biogrid Coherence KEGG", "String Coherence KEGG", "String Filt Coherence KEGG"))
# Remove large networks
mtx0_cor <- mtx0_cor[mtx0_cor$`Number of genes` < max_genes, ]
# Go through each pairs of columns and calculate pairwise complete observations
# pairwise_columns <- combn(ncol(mtx0_cor), 2)
# for(j in 1:ncol(pairwise_columns)) {
#   print(paste0(colnames(mtx0_cor)[pairwise_columns[1, j]], " - ", colnames(mtx0_cor)[pairwise_columns[2, j]], " : ", nrow(mtx0_cor[complete.cases(mtx0_cor[, c(pairwise_columns[1, j], pairwise_columns[2, j])]), c(pairwise_columns[1, j], pairwise_columns[2, j])])))
# }
# Actual correlation
diff_cors <- cor(mtx0_cor, use = "pairwise.complete.obs")

# corrgram::corrgram(diff_cors, order = FALSE, upper.panel=panel.cor)
corrplot.mixed(diff_cors, lower = "number", lower.col = col3(100), upper = "circle", upper.col = col3(100), tl.cex = 1, tl.pos = "lt", tl.col = "black")
# Selected correlations
png(filename = "Figures/FigS1_corr_coherence.png", height = 2000, width = 2000, res = 300)
# corrgram::corrgram(diff_cors, order = FALSE, upper.panel=panel.cor, cex.labels = 4)
corrplot.mixed(diff_cors, lower = "number", lower.col = col3(100), upper = "circle", upper.col = col3(100), tl.cex = 1, tl.pos = "lt", tl.col = "black")
dev.off()
```

### FigS2_corr_references.png

**Supplementary Figure S2.[FILE: Figures/FigS2_corr_references.png] Coherence estimates are similar when using different references of high coherences and PPI databases.** Red/blue gradient and numbers represent Pearson correlation coefficients between corresponding entries.

```{r}
# Selected data for correlation
# "Disease", "Category", "Number of SNPs", "Number of genes", "Biogrid Coherence", "String Coherence", "String Filt Coherence", "Biogrid Coherence Reactome", "String Coherence Reactome", "String Filt Coherence Reactome""Biogrid Coherence CC", "String Coherence CC", "String Filt Coherence CC", "Biogrid P-value", "String P-value", "String Filt P-value" 
mtx0_cor <- mtx0 %>% dplyr::select(c("Biogrid Coherence KEGG", "String Coherence KEGG", "String Filt Coherence KEGG", "Biogrid Coherence Reactome", "String Coherence Reactome", "String Filt Coherence Reactome", "Biogrid Coherence GOCC", "String Coherence GOCC", "String Filt Coherence GOCC"))
# Remove large networks
# mtx0_cor <- mtx0_cor[mtx0_cor$`Number of genes` < max_genes, ]
# Go through each pairs of columns and calculate pairwise complete observations
# pairwise_columns <- combn(ncol(mtx0_cor), 2)
# for(j in 1:ncol(pairwise_columns)) {
#   print(paste0(colnames(mtx0_cor)[pairwise_columns[1, j]], " - ", colnames(mtx0_cor)[pairwise_columns[2, j]], " : ", nrow(mtx0_cor[complete.cases(mtx0_cor[, c(pairwise_columns[1, j], pairwise_columns[2, j])]), c(pairwise_columns[1, j], pairwise_columns[2, j])])))
# }
# Actual correlation
diff_cors <- cor(mtx0_cor, use = "pairwise.complete.obs")

paste0("Average Pearson: ", diff_cors[upper.tri(diff_cors)] %>% mean()) # Average correlation

# corrgram::corrgram(diff_cors, order = FALSE, upper.panel=panel.cor)
corrplot.mixed(diff_cors, lower = "number", lower.col = col3(100), upper = "circle", upper.col = col3(100), tl.cex = 1, tl.pos = "lt", tl.col = "black")
# Selected correlations
png(filename = "Figures/FigS2_corr_references.png", height = 2000, width = 2000, res = 300)
# corrgram::corrgram(diff_cors, order = FALSE, upper.panel=panel.cor, cex.labels = 4)
corrplot.mixed(diff_cors, lower = "number", lower.col = col3(100), upper = "circle", upper.col = col3(100), tl.cex = 1, tl.pos = "lt", tl.col = "black")
dev.off()
```

\pagebreak

# Summary plots

```{r}
mtx_traits_selected_dotplot <- function(category = "Intelligence", mtx_traits = mtx_traits, databases = c("Biogrid", "STRING full", "STRING filtered"), reference = c("KEGG")) {
  mtx_traits_selected <- mtx_traits[mtx_traits$Category == category, grepl(reference, colnames(mtx_traits))]

  mtx_traits_selected_Biogrid <- data.frame(Disease = rownames(mtx_traits_selected), 
                                            Pvalue = mtx_traits_selected[, grepl("P-value", colnames(mtx_traits_selected)) & grepl("Biogrid", colnames(mtx_traits_selected))], 
                                            Coherence = mtx_traits_selected[, grepl("Coherence", colnames(mtx_traits_selected)) & grepl("Biogrid", colnames(mtx_traits_selected))], 
                                            PPI = paste0("Biogrid ", reference))

  mtx_traits_selected_STRING_full <- data.frame(Disease = rownames(mtx_traits_selected), 
                                            Pvalue = mtx_traits_selected[, grepl("P-value", colnames(mtx_traits_selected)) & grepl("String", colnames(mtx_traits_selected)) & !grepl("Filt", colnames(mtx_traits_selected))], 
                                            Coherence = mtx_traits_selected[, grepl("Coherence", colnames(mtx_traits_selected)) & grepl("String", colnames(mtx_traits_selected)) & !grepl("Filt", colnames(mtx_traits_selected))], 
                                            PPI = paste0("STRING full ", reference))

  mtx_traits_selected_STRING_filt <- data.frame(Disease = rownames(mtx_traits_selected), 
                                            Pvalue = mtx_traits_selected[, grepl("P-value", colnames(mtx_traits_selected)) & grepl("String Filt", colnames(mtx_traits_selected))], 
                                            Coherence = mtx_traits_selected[, grepl("Coherence", colnames(mtx_traits_selected)) & grepl("String Filt", colnames(mtx_traits_selected))], 
                                            PPI = paste0("STRING filtered ", reference))
  # Append plots from selected databases
  mtx_traits_selected_to_plot <- rbind()
  if( "Biogrid" %in% databases) {
    mtx_traits_selected_to_plot <- rbind(mtx_traits_selected_to_plot, mtx_traits_selected_Biogrid)
  }
  if( "STRING full" %in% databases) {
    mtx_traits_selected_to_plot <- rbind(mtx_traits_selected_to_plot, mtx_traits_selected_STRING_full)
  }
  if( "STRING filtered" %in% databases) {
    mtx_traits_selected_to_plot <- rbind(mtx_traits_selected_to_plot, mtx_traits_selected_STRING_filt)
  }

  mtx_traits_selected_to_plot <- mtx_traits_selected_to_plot[complete.cases(mtx_traits_selected_to_plot), ]

  Coherence_levels <- mtx_traits_selected_to_plot[mtx_traits_selected_to_plot$PPI == paste0("STRING filtered ", reference), ]
  if (length(setdiff(mtx_traits_selected_to_plot$Disease, Coherence_levels$Disease)) > 0) {
    Coherence_levels <- rbind(Coherence_levels, mtx_traits_selected_to_plot[(mtx_traits_selected_to_plot$Disease %in% setdiff(mtx_traits_selected_to_plot$Disease, Coherence_levels$Disease)) & mtx_traits_selected_to_plot$PPI != paste0("STRING filtered ", reference), , drop = FALSE])
  }

  mtx_traits_selected_to_plot$Disease <- factor(mtx_traits_selected_to_plot$Disease, levels = Coherence_levels$Disease[order(Coherence_levels$Coherence)])
  PPI_levels <- unique(mtx_traits_selected_to_plot$PPI)
  mtx_traits_selected_to_plot$PPI <- factor(mtx_traits_selected_to_plot$PPI, levels = c(PPI_levels[grepl("filtered", PPI_levels)], PPI_levels[grepl("full", PPI_levels)], PPI_levels[grepl("Biogrid", PPI_levels)]))
  # mtx_traits_selected_to_plot$Pvalue <- -log10(mtx_traits_selected_to_plot$Pvalue)

  p <- ggplot(data=mtx_traits_selected_to_plot, aes(x=Disease, y=Coherence, colour=PPI, size = Coherence)) + # label = Pvalue, 
    geom_point() +
    facet_grid(. ~ PPI) + 
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    ylab("Normalized coherence") +
    # geom_text_repel(size = 4, colour = "black") +
    coord_flip() +
    # theme_cowplot(12) +
    theme(legend.position="none") +
    theme(axis.text = element_text(size = 12)) + 
    ggtitle(category)
}
```

## Traits

### Fig3_summary_traits.png

```{r results='hide'}
# STRING filtered plots only
p_intelligence <- plot(mtx_traits_selected_dotplot(category = "Intelligence", mtx_traits = mtx0, databases = "STRING filtered", reference = "KEGG"))
p_anthropometric <- plot(mtx_traits_selected_dotplot(category = "Anthropometric trait", mtx_traits = mtx0, databases = "STRING filtered", reference = "KEGG"))
p_cardiovascular <- plot(mtx_traits_selected_dotplot(category = "Cardiovascular trait", mtx_traits = mtx0, databases = "STRING filtered", reference = "KEGG"))
p_eye <- plot(mtx_traits_selected_dotplot(category = "Eye trait", mtx_traits = mtx0, databases = "STRING filtered", reference = "KEGG"))
p_metabolic <- plot(mtx_traits_selected_dotplot(category = "Metabolic trait", mtx_traits = mtx0, databases = "STRING filtered", reference = "KEGG"))

png(filename = "Figures/Fig3_summary_traits.png", height = 2500, width = 5000, res = 300)
lay1 <- rbind(c(1, 2, 3))
lay2 <- rbind(c(1, 2, 3))
g1 <- arrangeGrob(grobs = list(p_anthropometric, p_cardiovascular, p_metabolic), layout_matrix = lay1)
g2 <- arrangeGrob(grobs = list(p_eye, p_intelligence), layout_matrix = lay2)
grid.arrange(g1, g2, ncol = 1)
dev.off()
```

### FigS3_summary_traits_KEGG.png

```{r }
p_intelligence <- plot(mtx_traits_selected_dotplot(category = "Intelligence", mtx_traits = mtx0, reference = "KEGG"))
p_anthropometric <- plot(mtx_traits_selected_dotplot(category = "Anthropometric trait", mtx_traits = mtx0, reference = "KEGG"))
p_cardiovascular <- plot(mtx_traits_selected_dotplot(category = "Cardiovascular trait", mtx_traits = mtx0, reference = "KEGG"))
p_eye <- plot(mtx_traits_selected_dotplot(category = "Eye trait", mtx_traits = mtx0, reference = "KEGG"))
p_metabolic <- plot(mtx_traits_selected_dotplot(category = "Metabolic trait", mtx_traits = mtx0, reference = "KEGG"))

png(filename = "Figures/FigS3_summary_traits_KEGG.png", height = 2500, width = 6000, res = 300)
lay1 <- rbind(c(1, 2, 3))
lay2 <- rbind(c(1, 2, 3))
g1 <- arrangeGrob(grobs = list(p_anthropometric, p_cardiovascular, p_metabolic), layout_matrix = lay1)
g2 <- arrangeGrob(grobs = list(p_eye, p_intelligence), layout_matrix = lay2)
grid.arrange(g1, g2, ncol = 1)
dev.off()
```

### FigS4_summary_traits_GOCC.png

```{r }
p_intelligence <- plot(mtx_traits_selected_dotplot(category = "Intelligence", mtx_traits = mtx0, reference = "GOCC"))
p_anthropometric <- plot(mtx_traits_selected_dotplot(category = "Anthropometric trait", mtx_traits = mtx0, reference = "GOCC"))
p_cardiovascular <- plot(mtx_traits_selected_dotplot(category = "Cardiovascular trait", mtx_traits = mtx0, reference = "GOCC"))
p_eye <- plot(mtx_traits_selected_dotplot(category = "Eye trait", mtx_traits = mtx0, reference = "GOCC"))
p_metabolic <- plot(mtx_traits_selected_dotplot(category = "Metabolic trait", mtx_traits = mtx0, reference = "GOCC"))

png(filename = "Figures/FigS4_summary_traits_GOCC.png", height = 2500, width = 6000, res = 300)
lay1 <- rbind(c(1, 2, 3))
lay2 <- rbind(c(1, 2, 3))
g1 <- arrangeGrob(grobs = list(p_anthropometric, p_cardiovascular, p_metabolic), layout_matrix = lay1)
g2 <- arrangeGrob(grobs = list(p_eye, p_intelligence), layout_matrix = lay2)
grid.arrange(g1, g2, ncol = 1)
dev.off()
```

### FigS5_summary_traits_Reactome.png

```{r }
p_intelligence <- plot(mtx_traits_selected_dotplot(category = "Intelligence", mtx_traits = mtx0, reference = "Reactome"))
p_anthropometric <- plot(mtx_traits_selected_dotplot(category = "Anthropometric trait", mtx_traits = mtx0, reference = "Reactome"))
p_cardiovascular <- plot(mtx_traits_selected_dotplot(category = "Cardiovascular trait", mtx_traits = mtx0, reference = "Reactome"))
p_eye <- plot(mtx_traits_selected_dotplot(category = "Eye trait", mtx_traits = mtx0, reference = "Reactome"))
p_metabolic <- plot(mtx_traits_selected_dotplot(category = "Metabolic trait", mtx_traits = mtx0, reference = "Reactome"))

png(filename = "Figures/FigS5_summary_traits_Reactome.png", height = 2500, width = 6000, res = 300)
lay1 <- rbind(c(1, 2, 3))
lay2 <- rbind(c(1, 2, 3))
g1 <- arrangeGrob(grobs = list(p_anthropometric, p_cardiovascular, p_metabolic), layout_matrix = lay1)
g2 <- arrangeGrob(grobs = list(p_eye, p_intelligence), layout_matrix = lay2)
grid.arrange(g1, g2, ncol = 1)
dev.off()
```

## Diseases

### Fig4_summary_diseases.png

```{r results='hide'}
# STRING filtered plots only
p_intelligence <- plot(mtx_traits_selected_dotplot(category = "Autoimmune", mtx_traits = mtx0, databases = "STRING filtered", reference = "KEGG"))
p_anthropometric <- plot(mtx_traits_selected_dotplot(category = "Cancer", mtx_traits = mtx0, databases = "STRING filtered", reference = "KEGG"))
p_cardiovascular <- plot(mtx_traits_selected_dotplot(category = "Cardiovascular disease", mtx_traits = mtx0, databases = "STRING filtered", reference = "KEGG"))
p_eye <- plot(mtx_traits_selected_dotplot(category = "Eye disease", mtx_traits = mtx0, databases = "STRING filtered", reference = "KEGG"))
p_metabolic <- plot(mtx_traits_selected_dotplot(category = "Metabolic disease", mtx_traits = mtx0, databases = "STRING filtered", reference = "KEGG"))
p_neurologic <- plot(mtx_traits_selected_dotplot(category = "Neurologic", mtx_traits = mtx0, databases = "STRING filtered", reference = "KEGG"))
p_psychiatric <- plot(mtx_traits_selected_dotplot(category = "Psychiatric", mtx_traits = mtx0, databases = "STRING filtered", reference = "KEGG"))

png(filename = "Figures/Fig4_summary_diseases.png", height = 2500, width = 5000, res = 300)
lay1 <- rbind(c(1, 2, 3, 4))
lay2 <- rbind(c(1, 2, 3, 4))
g1 <- arrangeGrob(grobs = list(p_intelligence, p_anthropometric, p_neurologic, p_psychiatric), layout_matrix = lay1)
g2 <- arrangeGrob(grobs = list(p_cardiovascular, p_metabolic, p_eye), layout_matrix = lay2)
grid.arrange(g1, g2, ncol = 1)
dev.off()
```

### FigS6_summary_diseases_KEGG.png

```{r }
p_intelligence <- plot(mtx_traits_selected_dotplot(category = "Autoimmune", mtx_traits = mtx0, reference = "KEGG"))
p_anthropometric <- plot(mtx_traits_selected_dotplot(category = "Cancer", mtx_traits = mtx0, reference = "KEGG"))
p_cardiovascular <- plot(mtx_traits_selected_dotplot(category = "Cardiovascular disease", mtx_traits = mtx0, reference = "KEGG"))
p_eye <- plot(mtx_traits_selected_dotplot(category = "Eye disease", mtx_traits = mtx0, reference = "KEGG"))
p_metabolic <- plot(mtx_traits_selected_dotplot(category = "Metabolic disease", mtx_traits = mtx0, reference = "KEGG"))
p_neurologic <- plot(mtx_traits_selected_dotplot(category = "Neurologic", mtx_traits = mtx0, reference = "KEGG"))
p_psychiatric <- plot(mtx_traits_selected_dotplot(category = "Psychiatric", mtx_traits = mtx0, reference = "KEGG"))

png(filename = "Figures/FigS6_summary_diseases_KEGG.png", height = 3000, width = 8000, res = 300)
lay1 <- rbind(c(1, 2, 3, 4))
lay2 <- rbind(c(1, 2, 3, 4))
g1 <- arrangeGrob(grobs = list(p_intelligence, p_anthropometric, p_neurologic, p_psychiatric), layout_matrix = lay1)
g2 <- arrangeGrob(grobs = list(p_cardiovascular, p_metabolic, p_eye), layout_matrix = lay2)
grid.arrange(g1, g2, ncol = 1)
dev.off()
```

### FigS7_summary_diseases_GOCC.png

```{r }
p_intelligence <- plot(mtx_traits_selected_dotplot(category = "Autoimmune", mtx_traits = mtx0, reference = "GOCC"))
p_anthropometric <- plot(mtx_traits_selected_dotplot(category = "Cancer", mtx_traits = mtx0, reference = "GOCC"))
p_cardiovascular <- plot(mtx_traits_selected_dotplot(category = "Cardiovascular disease", mtx_traits = mtx0, reference = "GOCC"))
p_eye <- plot(mtx_traits_selected_dotplot(category = "Eye disease", mtx_traits = mtx0, reference = "GOCC"))
p_metabolic <- plot(mtx_traits_selected_dotplot(category = "Metabolic disease", mtx_traits = mtx0, reference = "GOCC"))
p_neurologic <- plot(mtx_traits_selected_dotplot(category = "Neurologic", mtx_traits = mtx0, reference = "GOCC"))
p_psychiatric <- plot(mtx_traits_selected_dotplot(category = "Psychiatric", mtx_traits = mtx0, reference = "GOCC"))

png(filename = "Figures/FigS7_summary_diseases_GOCC.png", height = 3000, width = 8000, res = 300)
lay1 <- rbind(c(1, 2, 3, 4))
lay2 <- rbind(c(1, 2, 3, 4))
g1 <- arrangeGrob(grobs = list(p_intelligence, p_anthropometric, p_neurologic, p_psychiatric), layout_matrix = lay1)
g2 <- arrangeGrob(grobs = list(p_cardiovascular, p_metabolic, p_eye), layout_matrix = lay2)
grid.arrange(g1, g2, ncol = 1)
dev.off()
```

### FigS8_summary_diseases_Reactome.png

```{r }
p_intelligence <- plot(mtx_traits_selected_dotplot(category = "Autoimmune", mtx_traits = mtx0, reference = "Reactome"))
p_anthropometric <- plot(mtx_traits_selected_dotplot(category = "Cancer", mtx_traits = mtx0, reference = "Reactome"))
p_cardiovascular <- plot(mtx_traits_selected_dotplot(category = "Cardiovascular disease", mtx_traits = mtx0, reference = "Reactome"))
p_eye <- plot(mtx_traits_selected_dotplot(category = "Eye disease", mtx_traits = mtx0, reference = "Reactome"))
p_metabolic <- plot(mtx_traits_selected_dotplot(category = "Metabolic disease", mtx_traits = mtx0, reference = "Reactome"))
p_neurologic <- plot(mtx_traits_selected_dotplot(category = "Neurologic", mtx_traits = mtx0, reference = "Reactome"))
p_psychiatric <- plot(mtx_traits_selected_dotplot(category = "Psychiatric", mtx_traits = mtx0, reference = "Reactome"))

png(filename = "Figures/FigS8_summary_diseases_Reactome.png", height = 3000, width = 8000, res = 300)
lay1 <- rbind(c(1, 2, 3, 4))
lay2 <- rbind(c(1, 2, 3, 4))
g1 <- arrangeGrob(grobs = list(p_intelligence, p_anthropometric, p_neurologic, p_psychiatric), layout_matrix = lay1)
g2 <- arrangeGrob(grobs = list(p_cardiovascular, p_metabolic, p_eye), layout_matrix = lay2)
grid.arrange(g1, g2, ncol = 1)
dev.off()
```

# Modularity investigation

```{r}
fileNameIn0  <- "tables/Mod_Table.xlsx"   # diseases and traits, coherence new KEGG, REACTOME, CC. 2019-07-14, Kellen
# iconv -f UTF-8 -t ISO-8859-1 -c Mod_Table.csv >tmp.txt
fileNameIn0  <- "tables/tmp.txt"
mtx0 <- read_csv(fileNameIn0) %>% as.data.frame()
rownames(mtx0) <- mtx0$Disease
```

```{r}
# Selected data for correlation
# "Disease", "Category", "Number of SNPs", "Number of genes", "Biogrid Coherence", "String Coherence", "String Filt Coherence", "Biogrid Coherence Reactome", "String Coherence Reactome", "String Filt Coherence Reactome""Biogrid Coherence CC", "String Coherence CC", "String Filt Coherence CC", "Biogrid P-value", "String P-value", "String Filt P-value" 
mtx0_cor <- mtx0 %>% dplyr::select(c("Number of SNPs", "Number of genes", "Biogrid Coherence KEGG", "String Coherence KEGG", "String Filt Coherence KEGG", "Modularity Biogrid", "Modularity String", "Modularity String Filt"))
# Remove large networks
mtx0_cor <- mtx0_cor[mtx0_cor$`Number of genes` < max_genes, ]
# Go through each pairs of columns and calculate pairwise complete observations
# pairwise_columns <- combn(ncol(mtx0_cor), 2)
# for(j in 1:ncol(pairwise_columns)) {
#   print(paste0(colnames(mtx0_cor)[pairwise_columns[1, j]], " - ", colnames(mtx0_cor)[pairwise_columns[2, j]], " : ", nrow(mtx0_cor[complete.cases(mtx0_cor[, c(pairwise_columns[1, j], pairwise_columns[2, j])]), c(pairwise_columns[1, j], pairwise_columns[2, j])])))
# }
# Actual correlation
diff_cors <- cor(mtx0_cor, use = "pairwise.complete.obs")

# corrgram::corrgram(diff_cors, order = FALSE, upper.panel=panel.cor)
corrplot.mixed(diff_cors, lower = "number", lower.col = col3(100), upper = "circle", upper.col = col3(100), tl.cex = 1, tl.pos = "lt", tl.col = "black")
# Selected correlations
png(filename = "Figures/FigS1_corr_coherence.png", height = 2000, width = 2000, res = 300)
# corrgram::corrgram(diff_cors, order = FALSE, upper.panel=panel.cor, cex.labels = 4)
corrplot.mixed(diff_cors, lower = "number", lower.col = col3(100), upper = "circle", upper.col = col3(100), tl.cex = 1, tl.pos = "lt", tl.col = "black")
dev.off()
```