### 2019-08-17 manuscript/misc/07-07-2019_Dozmorov_progress.Rmd

---
# All BibTex reference files
bibliography: [/Users/mdozmorov/Documents/Work/VCU_grants/1_Dozmorov/3D_refs.bib,/Users/mdozmorov/Documents/Work/VCU_grants/1_Dozmorov/brain.bib,/Users/mdozmorov/Documents/Work/VCU_grants/1_Dozmorov/BRCA.bib,/Users/mdozmorov/Documents/Work/VCU_grants/1_Dozmorov/education.bib,/Users/mdozmorov/Documents/Work/VCU_grants/1_Dozmorov/enrichment.bib,/Users/mdozmorov/Documents/Work/VCU_grants/1_Dozmorov/networks.bib,/Users/mdozmorov/Documents/Work/VCU_grants/1_Dozmorov/statistics.bib]
# Reference template
csl: /Users/mdozmorov/Documents/Work/presentations/Templates/grant_template/styles.ref/genomebiology.csl
# Output template
output: 
  word_document:
    reference_docx: /Users/mdozmorov/Documents/Work/presentations/Templates/grant_template/styles.doc/NIH_grant_style.docx
---

## Progress report, Dozmorov M., 07-07-2019

### Coherence of Molecular Mechanisms in Genes Identified Through GWAS In Major Human Disease and Traits

This project involves quantification of coherence of the molecular networks formed by disease-associated genes. A metric for coherence is defined as a ratio between the number of internal and external protein-protein interactions. To date, a major problem with network size affecting the coherence has been solved. The solution includes the creation of size-matched references for high and low coherences (KEGG and random networks, respectively). The project is currently in results-generating and manuscript writing stage.