### 2019-08-20 Analysis/Generate_KEGG.Rmd
finish kegg/react/cc generation

---
title: "Generate_KEGG"
author: "Kellen Cresswell"
date: "August 19, 2019"
output: pdf_document
---

```{r}
require(msigdbr)
source("./functions/disease2frame_raw.R")
```

```{r}
#Read in msig data
msig_full = msigdbr()

sig_sub = msig_full %>% filter(gs_subcat %in% c("CP:REACTOME", "CP:KEGG", "CC")) %>% 
    group_by(gs_name, gs_subcat) %>% mutate(Count = n()) %>% group_by(gs_name, gs_subcat, Count) %>%
    do({
      sub = diseases2frame_raw(.$gene_symbol, string, .$gs_name[1])
      mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = sub)
      data.frame(Slope= coef(mod)[1])
    })
```

```{r}

```