### 2019-08-27 Analysis/Coherence_Calculations_New.Rmd
fix paths in commented sections

---
title: "New_Data_Coherence"
author: "Kellen Cresswell"
date: "July 1, 2019"
output: pdf_document
---

```{r}
require(msigdbr)
require(dplyr)
require(readr)
source("./functions/GeneSample.R")
require(biomaRt)

```

```{r}
# #Repeating for string
# 
# data_dir = "./data/"
# 
# #Read in new string database
# 
# string = read_table2(paste0(data_dir,"9606.protein.links.v11.0.txt.gz"))
# 
# string = string %>% mutate(protein1 = gsub("9606.", "", protein1), 
#                              protein2 = gsub("9606.", "", protein2))
# 
# #Read in ensemble protein ID
# 
# ensembl = biomaRt::useEnsembl(biomart="ensembl", dataset="hsapiens_gene_ensembl", mirror = "www")
# 
# #Get gene set
# genes = biomaRt::getBM(attributes=c('ensembl_peptide_id','hgnc_symbol'), mart = ensembl)
# 
# 
# #Join gene names with string
# 
# string = left_join(string, genes, by = c("protein1" = "ensembl_peptide_id", "protein2" = "ensembl_peptide_id"))
# 
# #Rename to work with our functions
# 
# string = string %>% dplyr::rename(hgnc_symbol_a = hgnc_symbol)
# 
# #Repeat for second set of proteins
# 
# string = left_join(string, genes, by = c("protein2" = "ensembl_peptide_id"))
# 
# string = string %>% dplyr::rename(hgnc_symbol_b = hgnc_symbol)
# 
# string_filt = string %>% filter(combined_score>=500)
# 
# dim(string)
# 
```

```{r}
# #Reading in biogrid data
# 
# links = read_tsv("./data/BIOGRID-ALL-3.5.174.mitab.txt")
# 
# #Select human protein-protein interactions
# 
# links = links %>% filter(`Taxid Interactor A` == "taxid:9606" & `Taxid Interactor B` == "taxid:9606")
# 
# #Split based on Biogrid ID system then pull out gene names
# 
# protein1 = lapply(strsplit(links$`Alt IDs Interactor A`, "locuslink:"), function(x) strsplit(x[2], "\\|")[[1]][1])
# protein2 = lapply(strsplit(links$`Alt IDs Interactor B`, "locuslink:"), function(x) strsplit(x[2], "\\|")[[1]][1])
# 
# protein1 = unlist(protein1)
# protein2 = unlist(protein2)
# 
# #Place into matrix in form that disease2frame works on
# 
# links = links %>% mutate(protein1 =protein1,
#                              protein2 = protein2) %>% dplyr::select(protein1, protein2)
# 
# links = links %>% dplyr::select(hgnc_symbol_a = protein1,hgnc_symbol_b = protein2)
# 
# 
# dim(links)

```

```{r}
#Read in edges generated from Connectivity_Comparison_*.Rmd

disease_overall_string = readRDS("string_edges_new.rds")

disease_overall_string_filt = readRDS("string_filt_edges_new.rds")

disease_overall_biogrid = readRDS("biogrid_edges_new.rds")

#Calculate slopes for all diseases under each category

dis_slopes_string_count = disease_overall_string %>% filter(Internal != 0) %>% filter(Category != "Random") %>% filter(Category != "KEGG") %>%
    group_by(Disease, Category) %>% mutate(Count = n()) %>% group_by(Disease, Category, Count) %>%
    do({
      mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = .)
      data.frame(Slope= coef(mod)[1])
    }) %>% arrange(Slope) %>% filter(Count>4)

dis_slopes_string_filt_count = disease_overall_string_filt %>% filter(Internal != 0) %>% filter(Category != "Random") %>% filter(Category != "KEGG") %>%
    group_by(Disease, Category) %>% mutate(Count = n()) %>% group_by(Disease, Category, Count) %>%
    do({
      mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = .)
      data.frame(Slope= coef(mod)[1])
    }) %>% arrange(Slope) %>% filter(Count>4)

dis_slopes_biogrid_count = disease_overall_biogrid %>% filter(Internal != 0) %>% filter(Category != "Random") %>% filter(Category != "KEGG") %>%
    group_by(Disease, Category) %>% mutate(Count = n()) %>% group_by(Disease, Category, Count) %>%
    do({
      mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = .)
      data.frame(Slope= coef(mod)[1])
    }) %>% arrange(Slope) %>% filter(Count>4)

# #Running coherence for string
# 
# #Read in KEGG pathways for coherence
# msig = readRDS("KEGG_String.rds")
# msig =msig %>% filter(Category == "CP:KEGG")
# coherence_tab = bind_rows()
# for (i in unique(dis_slopes_string_count$Disease)) {
#   #Filter to only include specific disease coherence
#   dis = dis_slopes_string_count %>% ungroup() %>% filter(Disease == i)
#   #Get closest KEGG pathways in term of gene counts
#   msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
#   #Take top 10 KEGG pathways
#   msigs = msigs[msigs$Num_Genes %in% unique(msigs$Num_Genes)[1:10],]
#   
#   diseases = bind_rows()
#   #Generate random networks
#   j = 1
#     while(j<11) {
#     #Sample gene networks equal in size to disease count
#     diseases_curr = GeneSample(dis$Count, string, directed = FALSE)
#     #Get random gene networks
#     diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
#     
#     #Repeat if there are no internal edges
#     if (any(diseases_curr$Internal !=0)) {
#       j = j+1
#       diseases = bind_rows(diseases, diseases_curr)
# 
#     }
#     }
#     #Calculate overall slope of random disease networks
#       mod_rand = lm(External ~ Internal-1, data = diseases)
#       rand_slope = coef(mod_rand)[[1]]
#       
#       #Normalize based on random slope and the median of kegg slopes (More robust)
#       coherence = (dis$Slope - rand_slope)/(median(msigs$Slope, na.rm = TRUE)-rand_slope)
#       coherence = data.frame(Disease = dis$Disease, Coherence = coherence, Msig_Norm = median(msigs$Slope, na.rm = TRUE), Rand_Norm = rand_slope )      
#       coherence_tab = bind_rows(coherence, coherence_tab)
#       saveRDS(coherence_tab, "./data/Coherence_Results/coherence_string.rds")
# }
#   
# 
# ```

#STRING All Msig

# ```{r}
# for (j in unique(msig$gs_name)) {
#   curr_msig = msig %>% filter(gs_name == j)
#   write.table(curr_msig$gene_symbol, file = "temp_rea_string.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
#   diseases_curr = diseases2frame("temp_rea_string.txt", links = string)
#   if (nrow(diseases_curr) == 0) {
#     next
#   }
#   mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = diseases_curr)
#   mod_slope = coef(mod)[[1]]
#   mod_sum = data.frame(Disease = j, Count =nrow(diseases_curr), Slope = mod_slope,
#                        Category = curr_msig$gs_subcat[1])
#   mod_over = bind_rows(mod_over, mod_sum)
#   saveRDS(mod_over,"./data/Coherence_Results/All_Msig_String.rds")
# }
# ```

# ```{r}
# msig = readRDS("KEGG_String_Filt.rds")
# msig =msig %>% filter(Category == "CP:KEGG")
# coherence_tab = bind_rows()
# for (i in unique(dis_slopes_string_filt_count$Disease)) {
#   print(i)
#   dis = dis_slopes_string_filt_count %>% filter(Disease == i)
#   msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
#   msigs = msigs[msigs$Num_Genes %in% unique(msigs$Num_Genes)[1:10],]
# 
#   diseases = bind_rows()
#   j = 1
#     while(j<11) {
#     diseases_curr = GeneSample(dis$Count, string_filt, directed = FALSE)
#     diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
#     if (any(diseases_curr$Internal !=0)) {
#       j = j+1
#       diseases = bind_rows(diseases, diseases_curr)
# 
#     }
#     }
# 
#       mod_rand = lm(External ~ Internal-1, data = diseases)
#       rand_slope = coef(mod_rand)[[1]]
# 
#       coherence = (dis$Slope - rand_slope)/(median(msigs$Slope, na.rm = TRUE)-rand_slope)
#     coherence = data.frame(Disease = dis$Disease, Coherence = coherence, Msig_Norm = median(msigs$Slope, na.rm = TRUE), Rand_Norm = rand_slope )
#       coherence_tab = bind_rows(coherence, coherence_tab)
#       saveRDS(coherence_tab, "./data/Coherence_Results/coherence_string_filt.rds")
# }

```

```{r}
# msig = readRDS("KEGG_Biogrid.rds")
# msig =msig %>% filter(Category == "CP:KEGG")
# coherence_tab = bind_rows()
# for (i in unique(dis_slopes_biogrid_count$Disease)) {
#   print(i)
#   dis = dis_slopes_biogrid_count %>% filter(Disease == i)
#   msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
#   msigs = msigs[msigs$Num_Genes %in% unique(msigs$Num_Genes)[1:10],]
# 
#   diseases = bind_rows()
#   j = 1
#     while(j<11) {
#     diseases_curr = GeneSample(dis$Count, links, directed = FALSE)
#     diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
#     if (any(diseases_curr$Internal !=0)) {
#       j = j+1
#       diseases = bind_rows(diseases, diseases_curr)
# 
#     }
#     }
# 
#       mod_rand = lm(External ~ Internal-1, data = diseases)
#       rand_slope = coef(mod_rand)[[1]]
# 
#       coherence = (dis$Slope - rand_slope)/(median(msigs$Slope, na.rm = TRUE)-rand_slope)
#       coherence = data.frame(Disease = dis$Disease, Coherence = coherence, Msig_Norm = median(msigs$Slope, na.rm = TRUE), Rand_Norm = rand_slope )
#       coherence_tab = bind_rows(coherence, coherence_tab)
#       print(coherence_tab)
#       saveRDS(coherence_tab, "./data/Coherence_Results/coherence_biogrid.rds")
# }

```

# Reactome

```{r}
# msig = readRDS("KEGG_String.rds")
# msig =msig %>% filter(Category == "CP:REACTOME") %>% filter(!is.na(Slope))
# coherence_tab = bind_rows()
# for (i in unique(dis_slopes_string_count$Disease)) {
#   print(i)
#   dis = dis_slopes_string_count %>% filter(Disease == i)
#   msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
#   msigs = msigs[msigs$Num_Genes %in% unique(msigs$Num_Genes)[1:10],]
#   
#   diseases = bind_rows()
#   j = 1
#     while(j<11) {
#     diseases_curr = GeneSample(dis$Count, string, directed = FALSE)
#     diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
#     if (any(diseases_curr$Internal !=0)) {
#       j = j+1
#       diseases = bind_rows(diseases, diseases_curr)
# 
#     }
#     }
#   
#       mod_rand = lm(External ~ Internal-1, data = diseases)
#       rand_slope = coef(mod_rand)[[1]]
#       
#       coherence = (dis$Slope - rand_slope)/(median(msigs$Slope, na.rm = TRUE)-rand_slope)
#       coherence = data.frame(Disease = dis$Disease, Coherence = coherence, Msig_Norm = median(msigs$Slope, na.rm = TRUE), Rand_Norm = rand_slope )
#       
#       coherence_tab = bind_rows(coherence, coherence_tab)
#       saveRDS(coherence_tab, "./data/Coherence_Results/coherence_string_react.rds")
# }
#   
# 
#
```

```{r}
# msig = readRDS("KEGG_String_Filt.rds")
# msig =msig %>% filter(Category == "CP:REACTOME") %>% filter(!is.na(Slope))
# coherence_tab = bind_rows()
# for (i in unique(dis_slopes_string_filt_count$Disease)) {
#   print(i)
#   dis = dis_slopes_string_filt_count %>% filter(Disease == i)
#   msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
#   msigs = msigs[msigs$Num_Genes %in% unique(msigs$Num_Genes)[1:10],]
#   
#   diseases = bind_rows()
#   j = 1
#     while(j<11) {
#     diseases_curr = GeneSample(dis$Count, string_filt, directed = FALSE)
#     diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
#     if (any(diseases_curr$Internal !=0)) {
#       j = j+1
#       diseases = bind_rows(diseases, diseases_curr)
# 
#     }
#     }
#   
#       mod_rand = lm(External ~ Internal-1, data = diseases)
#       rand_slope = coef(mod_rand)[[1]]
#       
#       coherence = (dis$Slope - rand_slope)/(median(msigs$Slope, na.rm = TRUE)-rand_slope)
#       coherence = data.frame(Disease = dis$Disease, Coherence = coherence)
#       
#       coherence_tab = bind_rows(coherence, coherence_tab)
#       saveRDS(coherence_tab, "./data/Coherence_Results/coherence_string_filt_react.rds")
# }
#   
#   
# 
```

```{r}
# msig = readRDS("KEGG_Biogrid.rds")
# msig =msig %>% filter(Category == "CP:REACTOME") %>% filter(!is.na(Slope))
# coherence_tab = bind_rows()
# for (i in unique(dis_slopes_biogrid_count$Disease)) {
#   print(i)
#   dis = dis_slopes_biogrid_count %>% filter(Disease == i)
#   msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
#   msigs = msigs[msigs$Num_Genes %in% unique(msigs$Num_Genes)[1:10],]
# 
#   diseases = bind_rows()
#   j = 1
#     while(j<11) {
#     diseases_curr = GeneSample(dis$Count, links, directed = FALSE)
#     diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
#     if (any(diseases_curr$Internal !=0)) {
#       j = j+1
#       diseases = bind_rows(diseases, diseases_curr)
# 
#     }
#     }
# 
#       mod_rand = lm(External ~ Internal-1, data = diseases)
#       rand_slope = coef(mod_rand)[[1]]
# 
#       coherence = (dis$Slope - rand_slope)/(median(msigs$Slope, na.rm = TRUE)-rand_slope)
#       coherence = data.frame(Disease = dis$Disease, Coherence = coherence)
# 
#       coherence_tab = bind_rows(coherence, coherence_tab)
#       saveRDS(coherence_tab, "./data/Coherence_Results/coherence_biogrid_react.rds")
# }
# 
# 
```

# CC

```{r}
# msig = readRDS("KEGG_String.rds")
# msig =msig %>% filter(Category == "CC") %>% filter(!is.na(Slope))
# coherence_tab = bind_rows()
# for (i in unique(dis_slopes_string_count$Disease)) {
#   print(i)
#   dis = dis_slopes_string_count %>% filter(Disease == i)
#   msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
#   msigs = msigs[msigs$Num_Genes %in% unique(msigs$Num_Genes)[1:10],]
#   
#   diseases = bind_rows()
#   j = 1
#     while(j<11) {
#     diseases_curr = GeneSample(dis$Count, string, directed = FALSE)
#     diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
#     if (any(diseases_curr$Internal !=0)) {
#       j = j+1
#       diseases = bind_rows(diseases, diseases_curr)
# 
#     }
#     }
#   
#       mod_rand = lm(External ~ Internal-1, data = diseases)
#       rand_slope = coef(mod_rand)[[1]]
#       
#       coherence = (dis$Slope - rand_slope)/(median(msigs$Slope, na.rm = TRUE)-rand_slope)
#       coherence = data.frame(Disease = dis$Disease, Coherence = coherence, Msig_Norm = median(msigs$Slope, na.rm = TRUE), Rand_Norm = rand_slope )
#       
#       coherence_tab = bind_rows(coherence, coherence_tab)
#       saveRDS(coherence_tab, "./data/Coherence_Results/coherence_string_CC.rds")
# }
#   
# 
# ```
# 
# ```{r}
# msig = readRDS("KEGG_String_Filt.rds")
# msig =msig %>% filter(Category == "CC") %>% filter(!is.na(Slope))
# coherence_tab = bind_rows()
# for (i in unique(dis_slopes_string_filt_count$Disease)) {
#   print(i)
#   dis = dis_slopes_string_filt_count %>% filter(Disease == i)
#   msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
#   msigs = msigs[msigs$Num_Genes %in% unique(msigs$Num_Genes)[1:10],]
# 
#   diseases = bind_rows()
#   j = 1
#     while(j<11) {
#     diseases_curr = GeneSample(dis$Count, string_filt, directed = FALSE)
#     diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
#     if (any(diseases_curr$Internal !=0)) {
#       j = j+1
#       diseases = bind_rows(diseases, diseases_curr)
# 
#     }
#     }
# 
#       mod_rand = lm(External ~ Internal-1, data = diseases)
#       rand_slope = coef(mod_rand)[[1]]
# 
#       coherence = (dis$Slope - rand_slope)/(median(msigs$Slope, na.rm = TRUE)-rand_slope)
#       coherence = data.frame(Disease = dis$Disease, Coherence = coherence)
# 
#       coherence_tab = bind_rows(coherence, coherence_tab)
#       saveRDS(coherence_tab, "./data/Coherence_Results/coherence_string_filt_CC.rds")
# }
#

```

```{r}
# msig = readRDS("KEGG_Biogrid.rds")
# msig =msig %>% filter(Category == "CC") %>% filter(!is.na(Slope))
# coherence_tab = bind_rows()
# for (i in unique(dis_slopes_biogrid_count$Disease)) {
#   print(i)
#   dis = dis_slopes_biogrid_count %>% filter(Disease == i)
#   msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
#   msigs = msigs[msigs$Num_Genes %in% unique(msigs$Num_Genes)[1:10],]
# 
#   diseases = bind_rows()
#   j = 1
#     while(j<11) {
#     diseases_curr = GeneSample(dis$Count, links, directed = FALSE)
#     diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
#     if (any(diseases_curr$Internal !=0)) {
#       j = j+1
#       diseases = bind_rows(diseases, diseases_curr)
# 
#     }
#     }
# 
#       mod_rand = lm(External ~ Internal-1, data = diseases)
#       rand_slope = coef(mod_rand)[[1]]
# 
#       coherence = (dis$Slope - rand_slope)/(median(msigs$Slope, na.rm = TRUE)-rand_slope)
#       coherence = data.frame(Disease = dis$Disease, Coherence = coherence)
# 
#       coherence_tab = bind_rows(coherence, coherence_tab)
#       saveRDS(coherence_tab, "./data/Coherence_Results/coherence_biogrid_CC.rds")
# }
```

```{r}
#Random networks

# #Read in KEGG pathways for coherence
# msig = readRDS("KEGG_String.rds")
# msig = msig %>% filter(Category == "CP:KEGG")
# string_random = bind_rows()
# for (i in msig$Count) {   
#   j = 0
#   diseases = bind_rows()
#     while(j<11) {
#     diseases_curr = GeneSample(i, string, directed = FALSE)
#     diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
#     if (any(diseases_curr$Internal !=0)) {
#       j = j+1
#       diseases = bind_rows(diseases, diseases_curr)
# 
#     }
#     }
#   
#       mod_rand = lm(External ~ Internal-1, data = diseases)
#       diseases_curr = data.frame(Count = i, Slope = coef(mod_rand)[1])
#       string_random = bind_rows(string_random, diseases_curr)
#       print(string_random)
#       saveRDS(string_random, "./data/Coherence_Results/string_random_edges.rds")
# }

```

```{r}
#Read in results generated above

#KEGG

coherence_string_filt = readRDS("./data/Coherence_Results/coherence_string_filt.rds") %>% dplyr::select(Disease, Coherence)
coherence_string = readRDS("./data/Coherence_Results/coherence_string.rds") %>% dplyr::select(Disease, Coherence)
coherence_biogrid = readRDS("./data/Coherence_Results/coherence_biogrid.rds") %>% dplyr::select(Disease, Coherence)

#Reactome

coherence_string_filt_react = readRDS("./data/Coherence_Results/coherence_string_filt_react.rds") %>% dplyr::select(Disease, Coherence)
coherence_string_react = readRDS("./data/Coherence_Results/coherence_string_react.rds") %>% dplyr::select(Disease, Coherence)
coherence_biogrid_react = readRDS("./data/Coherence_Results/coherence_biogrid_react.rds") %>% dplyr::select(Disease, Coherence)

#CC

coherence_string_filt_cc = readRDS("./data/Coherence_Results/coherence_string_filt_CC.rds") %>% dplyr::select(Disease, Coherence)
coherence_string_cc = readRDS("./data/Coherence_Results/coherence_string_CC.rds") %>% dplyr::select(Disease, Coherence)
coherence_biogrid_cc = readRDS("./data/Coherence_Results/coherence_biogrid_CC.rds") %>% dplyr::select(Disease, Coherence)

#Full join KEGG coherence results together based on disease
coherence_full = full_join(coherence_biogrid, coherence_string, by = "Disease")

coherence_full = full_join(coherence_full, coherence_string_filt, by = "Disease")

#Give columns proper name

colnames(coherence_full) = c("Disease", "Biogrid_Coherence", "String_Coherence", "String_Filt_Coherence")

#Save coherence full in this state for later manipulation
coherence_disease = coherence_full

```

```{r}
#Finish all disease plot (From save coherence disease frame)

#Joining REACTOME

coherence_disease = left_join(coherence_disease, coherence_biogrid_react, by= c("Disease"= "Disease"))

coherence_disease = left_join(coherence_disease, coherence_string_react, by= c("Disease"= "Disease"))

coherence_disease = left_join(coherence_disease, coherence_string_filt_react, by= c("Disease"= "Disease"))

#Joining CC

coherence_disease = left_join(coherence_disease, coherence_biogrid_cc, by= c("Disease"= "Disease"))

coherence_disease = left_join(coherence_disease, coherence_string_cc, by= c("Disease"= "Disease"))

coherence_disease = left_join(coherence_disease, coherence_string_filt_cc, by= c("Disease"= "Disease"))

coherence_disease = coherence_disease  %>% dplyr::rename(`Biogrid Coherence Reactome` = Coherence.x, `String Coherence Reactome` = Coherence.y, `String Filt Coherence Reactome` = Coherence.x.x, `Biogrid Coherence CC` = Coherence.y.y, `String Coherence CC` = Coherence.x.x.x, `String Filt Coherence CC` = Coherence.y.y.y)

#Get categories from slope frame

Cat = bind_rows(dis_slopes_string_count %>% ungroup() %>% dplyr::select(Disease, Category) %>% distinct(), dis_slopes_biogrid_count %>% ungroup() %>% dplyr::select(Disease, Category) %>% distinct())

#Join to all diseases frame

coherence_disease =left_join(coherence_disease, Cat)

#Remove underscores from column names
colnames(coherence_disease) = gsub("\\_", " ", colnames(coherence_disease))

coherence_disease = coherence_disease %>% dplyr::select(Disease, Category, `Biogrid Coherence`:`String Filt Coherence CC`) %>% dplyr::distinct()

#Adding gene counts

disease_select = read_csv("./data/supplementary_table_diseases_selected.csv") %>% dplyr::select(Disease =`Disease name`, `Number of SNPs`, `Number of genes`)

#Join gene counts

coherence_disease = left_join(coherence_disease, disease_select)

coherence_disease = coherence_disease %>% dplyr::select(Disease, Category, `Number of SNPs`, `Number of genes`, `Biogrid Coherence`:`String Filt Coherence CC`)
```

```{r}
#Reading in data from cluster and binding

biogrid_p = bind_rows(readRDS("./data/Permutation_Results/biogrid_p_1k.rds") %>% as.data.frame(.), readRDS("./data/Permutation_Results/biogrid_p_psych_1k.rds") %>% as.data.frame(.))

string_p = bind_rows(readRDS("./data/Permutation_Results/string_p_1k.rds") %>% as.data.frame(.), readRDS("./data/Permutation_Results/string_p_psych_1k.rds") %>% as.data.frame(.), readRDS("./data/Permutation_Results/string_p_neuro_1k.rds") %>% as.data.frame(.))

string_filt_p = bind_rows(readRDS("./data/Permutation_Results/string_filt_p_1k.rds") %>% as.data.frame(.), readRDS("./data/Permutation_Results/string_filt_p_psych_1k.rds") %>% as.data.frame(.), readRDS("./data/Permutation_Results/string_filt_p_neuro_1k.rds") %>% as.data.frame(.))

colnames(biogrid_p) = c("Disease", "Biogrid P-value", "Ave")
colnames(string_p) = c("Disease", "String P-value", "Ave")
colnames(string_filt_p) = c("Disease", "String Filt P-value", "Ave")

#Joining ignoring the average coherence column I saved

#Trait dataset 
coherence_full = left_join(coherence_full,biogrid_p %>% dplyr::select(-Ave))

coherence_full = left_join(coherence_full,string_p%>% dplyr::select(-Ave))

coherence_full = left_join(coherence_full,string_filt_p%>% dplyr::select(-Ave))

#Complete Disease dataset

coherence_disease = left_join(coherence_disease,biogrid_p %>% dplyr::select(-Ave), by = c("Disease" = "Disease"))

coherence_disease = left_join(coherence_disease,string_p %>% dplyr::select(-Ave), by = c("Disease" = "Disease"))

coherence_disease = left_join(coherence_disease,string_filt_p  %>% dplyr::select(-Ave), by = c("Disease" = "Disease"))

coherence_disease = coherence_disease %>% mutate_if(is.numeric, round, digits = 2)

coherence_disease = coherence_disease %>% mutate(`Biogrid P-value` = round( as.numeric(as.character(`Biogrid P-value`)),4)) %>% mutate(`String P-value` = round( as.numeric(as.character(`String P-value`)),4)) %>% mutate(`String Filt P-value` = round( as.numeric(as.character(`String Filt P-value`)),4))

coherence_disease =coherence_disease %>% arrange(-`String Coherence`)

#Rename KEGG pathways when necessary

coherence_disease = coherence_disease %>% dplyr::rename(`String Coherence KEGG` = `String Coherence`, `String Filt Coherence KEGG` = `String Filt Coherence`, `Biogrid Coherence KEGG` = `Biogrid Coherence`)

colnames(coherence_disease) = gsub("CC", "GOCC", colnames(coherence_disease))

#Writing to a tables

dir.create("../manuscript/tables")
write.csv(coherence_disease, ".././manuscript/tables/supplementary_table_S3.csv", 
          row.names = FALSE)

```