### 2019-08-27 Analysis/Connectivity_Comparison.Rmd

---
title: "Connectivity Comparison Undirected"
author: "Kellen Cresswell"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.height = 6, fig.width = 9, digits = 4)
```

```{r}
source("./functions/perm_slopes.R")
source("./functions/GeneSample.R")
source("./functions/genes2interactions.R")
source("./functions/disease2frame.R")
source("./functions/diseases2mod.R")

# install.packages("png")
# Download https://bioconductor.org/packages/release/bioc/src/contrib/KEGGREST_1.24.0.tar.gz
# Manually install "KEGGREST" from the file
library(KEGGREST) # BiocManager::install("KEGGREST")
library(KEGG.db)  # BiocManager::install("KEGG.db")
library(biomaRt)  # BiocManager::install("biomaRt")
library(readr)    # install.packages("readr")
library(dplyr)    # install.packages("dplyr")
library(ggplot2)  # install.packages("ggplot2")
library(broom)    # install.packages("broom")
library(stringr)  # install.packages("stringr")
library(tidyr)    # install.packages("tidyr")
library(igraph)   # install.packages("igraph")
```

```{r}
#Reading in biogrid data
links = read_tsv("./data/BIOGRID-ALL-3.5.174.mitab.txt.gz")
links = links %>% filter(`Taxid Interactor A` == "taxid:9606" & `Taxid Interactor B` == "taxid:9606")

#Split based on Biogrid ID system then pull out gene names
protein1 = lapply(strsplit(links$`Alt IDs Interactor A`, "locuslink:"), function(x) strsplit(x[2], "\\|")[[1]][1])
protein2 = lapply(strsplit(links$`Alt IDs Interactor B`, "locuslink:"), function(x) strsplit(x[2], "\\|")[[1]][1])
protein1 = unlist(protein1)
protein2 = unlist(protein2)

#Place into matrix in form that disease2frame works on
links = links %>% mutate(protein1 =protein1, protein2 = protein2) %>% dplyr::select(protein1, protein2)
links = links %>% dplyr::select(hgnc_symbol_a = protein1,hgnc_symbol_b = protein2)

dim(links)
```

```{r}
#Repeating for string
data_dir = "./data/"
string = read_table2(paste0(data_dir,"9606.protein.links.v11.0.txt.gz"))
string= string %>% mutate(protein1 = gsub("9606.", "", protein1), protein2 = gsub("9606.", "", protein2))

#Read in ensemble protein ID
ensembl = useEnsembl(biomart="ensembl", dataset="hsapiens_gene_ensembl", mirror = "www")
genes = biomaRt::getBM(attributes=c('ensembl_peptide_id','hgnc_symbol'), mart = ensembl)

#Join gene names with string
string = left_join(string, genes, by = c("protein1" = "ensembl_peptide_id", "protein2" = "ensembl_peptide_id"))
colnames(string)[4] ="hgnc_symbol_a"
string = left_join(string, genes, by = c("protein2" = "ensembl_peptide_id"))
colnames(string)[5] ="hgnc_symbol_b"
dim(string)
```

```{r}
#Get directory with diseases
disease_directory = "./data/Disease_Genes/"
disease_names = list.files(disease_directory)
disease_list = paste0("./data/Disease_Genes/", disease_names, "/")

#Read in manually selected diseases
disease_select = read_csv("./data/supplementary_table_diseases_selected.csv")
disease_traits = disease_select %>% dplyr::select(`Disease name`, Category)

#Pull out disease and category information from select diseases

#Change first column to make more clear
colnames(disease_select)[1] = "Disease"

#Filtering out NA in last row
disease_select = disease_select %>% filter(!is.na(Disease))

#Creating a vector of file names from disease select
disease_files =  unique(paste0(disease_select$Disease, ".txt"))

#Make full path
disease_selected = paste0("./data/Disease_Genes/", disease_files)
```

```{r}
#Loop through diseases and pull out internal and external edges
disease_length = 1:length(disease_selected)
disease_frames = lapply(disease_length, function(x) {

  #Run function to find internal and external connectivity 
  diseases = diseases2frame(disease_selected[x], links = links)

  #Record ppi missing
  per_no_ppi = 1- (nrow(diseases)/nrow(read.table(disease_selected[x])))
  return(list(diseases = diseases, lengths = per_no_ppi))
})

# #Calculate modularity
# 
# disease_mod = lapply(disease_length, function(x) {
# 
#   #Run function to find internal and external connectivity 
#   mod = diseases2mod(disease_selected[x], links = links)
#   
# })
# 
# saveRDS(disease_mod, "Modularity_Biogrid.rds")

#Pull out network information for overall data frame
disease_protein = lapply(disease_frames, function(x) x$diseases)

#Convert to data frame
disease_overall = bind_rows(disease_protein)
disease_overall = left_join(disease_overall, disease_select %>% dplyr::select(Disease, Category), "Disease")
saveRDS(disease_overall, "biogrid_edges_new.rds")

# #Add number of snps with no connection to original data frame
# 
# per_no_ppi_bio = sapply(disease_frames, function(x) x$lengths)
# 
# per_no_ppi_frame = data.frame(Disease = unique(disease_select$Disease), per_no_ppi_bio = per_no_ppi_bio)
# 
# disease_select = left_join(disease_select, per_no_ppi_frame, "Disease")
```

```{r}
#Repeat for string

#Loop through diseases and pull out internal and external edges
disease_frames_string = lapply(disease_length, function(x) {

  #Run function to find internal and external connectivity 
  diseases = diseases2frame(disease_selected[x], links = string)

  #Record ppi missing
  per_no_ppi = 1- (nrow(diseases)/nrow(read.table(disease_selected[x])))
  return(list(diseases = diseases, lengths = per_no_ppi))
})

#String

# disease_mod_string = lapply(disease_length, function(x) {
# 
#   #Run function to find internal and external connectivity 
#   mod = diseases2mod(disease_selected[x], links = string)
#   
# })
# 
# saveRDS(disease_mod_string, "Modularity_String.rds")

#Pull out network information for overall data frame
disease_protein_string = lapply(disease_frames_string, function(x) x$diseases)

#Convert to data frame
disease_overall_string = bind_rows(disease_protein_string)
disease_overall_string = left_join(disease_overall_string, disease_select %>% dplyr::select(Disease, Category), "Disease")

#Add number of snps with no connection to original data frame

# per_no_ppi_string = sapply(disease_frames_string, function(x) x$lengths)
# 
# per_no_ppi_frame = data.frame(Disease = unique(disease_select$Disease), per_no_ppi_string = per_no_ppi_string)
# 
# saveRDS(disease_overall_string, "string_edges_new.rds")
# 
# disease_select = left_join(disease_select, per_no_ppi_frame, "Disease")

saveRDS(disease_overall_string, "string_edges_new.rds")
```

```{r}
#Repeat for string filtered
string = string %>% filter(combined_score>=500)

#Diseases
disease_length = 1:length(disease_selected)

#Loop through diseases and pull out internal and external edges
disease_frames_string = lapply(disease_length, function(x) {

  #Run function to find internal and external connectivity 
  diseases = diseases2frame(disease_selected[x], links = string)

  #Record ppi missing
  per_no_ppi = 1- (nrow(diseases)/nrow(read.table(disease_selected[x])))
  return(list(diseases = diseases, lengths = per_no_ppi))
})

#Pull out network information for overall data frame
disease_protein_string = lapply(disease_frames_string, function(x) x$diseases)

#Convert to data frame
disease_overall_string = bind_rows(disease_protein_string)
disease_overall_string = left_join(disease_overall_string, disease_select %>% dplyr::select(Disease, Category), "Disease")

#disease_overall_string = #readRDS("disease_overall_string.rds")
saveRDS(disease_overall_string, "string_filt_edges_new.rds")
```