### 2019-09-14 Analysis/Normalized_Coherence.Rmd
put everything in functions

---
title: "Msigdf Biogrid"
author: "Kellen Cresswell"
date: "July 19, 2019"
output: pdf_document
---

# Read in functions

```{r}  
source("./functions/disease2frame.R")
source("./functions/GeneSample.R")
source("./functions/get_msig_slopes.R")

library(readr) #install.packages("readr")
library(dplyr) #install.packages("dplyr")
library(biomaRt) #BiocManager::install("biomaRt")
library(msigdbr) #BiocManager::install("msigdbr")
library(igraph) #install.packages("igraph")
library(tidyr) #install.packages("tidyr")
library(broom) #install.packages("broom")

```

# Setting options

```{r}
#Option for number of msig pathways to take from each category
num_msig = 10
#Option for number of random networks to generate (Will generate num_rand-1 networks)
num_rand = 1001
# Construct file name to save the results
run <- 1 # Placeholder for the run ID
fileNameOut <- file.path("./data/Coherence_Results/", paste0("coherence_msig_", num_msig, "_rand_", num_rand, "_", run, ".csv"))
```

# Read in biogrid data

```{r}
#Reading in biogrid data

links = read_tsv("./data/BIOGRID-ALL-3.5.174.mitab.txt.gz")

links = links %>% filter(`Taxid Interactor A` == "taxid:9606" & `Taxid Interactor B` == "taxid:9606")

#Split based on Biogrid ID system then pull out gene names

protein1 = lapply(strsplit(links$`Alt IDs Interactor A`, "locuslink:"), function(x) strsplit(x[2], "\\|")[[1]][1])
protein2 = lapply(strsplit(links$`Alt IDs Interactor B`, "locuslink:"), function(x) strsplit(x[2], "\\|")[[1]][1])

protein1 = unlist(protein1)
protein2 = unlist(protein2)

#Place into matrix in form that disease2frame works on

links = links %>% mutate(protein1 =protein1, 
                         protein2 = protein2) %>% dplyr::select(protein1, protein2)

links = links %>% dplyr::select(hgnc_symbol_a = protein1,hgnc_symbol_b = protein2)

```

# Read in string data and create the string filtered frame

```{r}
string = read_table2("./data/9606.protein.links.v11.0.txt.gz")

string = string %>% mutate(protein1 = gsub("9606.", "", protein1), 
                           protein2 = gsub("9606.", "", protein2))

#Read in ensemble protein ID

ensembl = useEnsembl(biomart="ensembl", dataset="hsapiens_gene_ensembl", mirror = "www")

genes = biomaRt::getBM(attributes=c('ensembl_peptide_id','hgnc_symbol'), mart = ensembl)

#Join gene names with string

string = left_join(string, genes, by = c("protein1" = "ensembl_peptide_id", "protein2" = "ensembl_peptide_id"))

string = left_join(string, genes, by = c("protein2" = "ensembl_peptide_id"))

string = string %>% dplyr::rename(hgnc_symbol_a = hgnc_symbol.x)

string = string %>% dplyr::rename(hgnc_symbol_b = hgnc_symbol.y)

string_filt = string %>% filter(combined_score>=500)
```

```{r}

get_msig_slopes = function(sub_cats, database) {
  require(msigdbr)
  msig = msigdbr()

  #Subset to include msig
  msig = msig %>% filter(gs_subcat %in% sub_cats)

  #Create empty list
  mod_over = list()
  #Set index for list entries
  index = 1
  #Loop through each pathway
  for (j in unique(msig$gs_name)) {
    #Subset to only include current pathway
    curr_msig = msig %>% filter(gs_name == j)
    #Write to table in format for disease2frame
    write.table(curr_msig$gene_symbol, file = "temp_rea_bio.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
    #Get internal/external connectivity
    diseases_curr = diseases2frame("temp_rea_bio.txt", links = links)
    #Get disease name
    diseases_curr = diseases_curr %>% mutate(Disease = j)
    mod_over[[index]] = diseases_curr
    index =  index+1
  } 
  #Bind together list

  mod_over = bind_rows(mod_over)

  #Applying categories

  msig_cats  = msig %>% 
    dplyr::select(Disease = gs_name, Category = gs_subcat) %>% 
    distinct() #Get categories in order

  #Joining categories

  mod_over = left_join(mod_over, msig_cats)

  #Getting gene counts

  mod_over = mod_over %>% group_by(Disease) %>% mutate(Num_Genes = n())
  return(mod_over)
}

```

```{r}
#Setting the three categories we want

subcats = c("CC", "CP:KEGG", "CP:REACTOME")

#Getting biogrid msig slopes

mod_over_biogrid = get_msig_slopes(subcats = subcats, database = links)

#Getting string msig slopes

mod_over_string= get_msig_slopes(subcats = subcats, database = string)

#Getting string filtered msig slopes

mod_over_string_filt = get_msig_slopes(subcats = subcats, database = string_filt)

#Saving the files

saveRDS(mod_over_biogrid, "Biogrid_Msig_Slopes.rds")
saveRDS(mod_over_string, "String_Msig_Slopes.rds")
saveRDS(mod_over_string_filt, "String_Filt_Msig_Slopes.rds")
```

#Define function for calculating slopes and counts from internal/external edges
```{r}
slopes_from_edges = function(edges) {

  edges %>% filter(Internal != 0) %>% filter(Category != "Random") %>% filter(Category != "KEGG") %>%
  group_by(Disease, Category) %>% mutate(Count = n()) %>% group_by(Disease, Category, Count) %>%
  do({
    mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = .)
    data.frame(Slope= coef(mod)[1])
  }) %>% arrange(Slope) %>% filter(Count>4)

}
```

# Get edges for diseases from Biogrid (Connectivity_Comparison.Rmd)

```{r}
#Biogrid

disease_overall_biogrid     <- readRDS("biogrid_edges_new.rds")

#Get Biogrid slopes and edge counts

dis_slopes_biogrid_count = slopes_from_edges(disease_overall_biogrid)

#String

disease_overall_string     <- readRDS("string_edges_new.rds")

#Get string slopes and edge counts

dis_slopes_string_count = slopes_from_edges(disease_overall_string)

#String Filtered

disease_overall_string_filt     <- readRDS("string_filt_edges_new.rds")

#Get string filtered slopes and edge counts

dis_slopes_string_filt_count = slopes_from_edges(disease_overall_string_filt)

```

# Define function for calculating normalized coherence and counts ignoring zero internal edges

```{r}
calc_norm_coherence = function(msig_frame, dis_frame, num_msig = 10, num_rand = 10) {
  Count_Sum = msig_frame %>% dplyr::filter(Internal>0) %>% 
  group_by(Disease) %>%
  mutate(Num_Genes = n()) %>% dplyr::select(Disease,Category, Num_Genes) %>% distinct()
#Splitting into list by categories
Count_Sum = split(Count_Sum, Count_Sum$Category)
coherence_tab = bind_rows()
for (i in unique(dis_frame$Disease)) {
  #Get disease specific genes
  dis = dis_frame %>% filter(Disease == i)
  #Get 30 closest msigdf networks taking 10 from each database and bind
  close_nets = bind_rows(lapply(Count_Sum,  function(x) x %>% mutate(Diff = abs(Num_Genes-dis$Count)) %>%
    arrange(Diff) %>% head(.,num_msig)))
  #Subsetting msig to only include close pathways
  msig_sub = msig_frame %>% filter(Disease %in% close_nets$Disease)
  #Taking slopes for each category
  mod_msig = msig_sub %>% group_by(Disease) %>% do(tidy(
    coef(lm(I(sqrt(External)) ~ I(sqrt(Internal))-1, data = .))[[1]]
  ))
  #Take median
  mod_slope = median(mod_msig$x)
  #Creating 10 random networks
  j = 1
  slope_list = list()
    while(j<num_rand) {
    #Empty data frame for disease networks
      diseases = bind_rows()
      #Generate random networks until there as many internal edges as current diseases
      while(nrow(diseases)<dis$Count) {
    #Generate random diseases network
    diseases_curr = GeneSample(dis$Count, links, directed = FALSE)
    #Save internal and external edges
    diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = diseases_curr$internal_connectivity, External = diseases_curr$external_connectivity)
    #Filter out zeros
    diseases_curr = diseases_curr %>% filter(Internal !=0)
    #Bind together non-zero internal/external edges
    diseases = bind_rows(diseases, diseases_curr)
      }
      #Save slope of current network
      slope_list[[j]] = coef(lm(I(sqrt(External)) ~ I(sqrt(Internal))-1, data = diseases))[[1]]
      #Iterate
      j = j+1

    }

      #Get median of all slopes
      rand_slope = median(unlist(slope_list))

      #Calculate coherence
      coherence = (dis$Slope - rand_slope)/(mod_slope-rand_slope)
      #Place into data frame with disease, coherence, msig normalization value and random slope normalization value
      coherence = data.frame(Disease = dis$Disease, Coherence = coherence, 
                             Msig_Norm = mod_slope, Rand_Norm = rand_slope, Count = dis$Count)
      coherence_tab = bind_rows(coherence, coherence_tab)
      return(coherence_tab)
}
}
```

# Calculating normalized coherence

```{r}
#Get read in msig pathway genes for KEGG, Reactome and GOCC
msig_frame_biogrid = readRDS("Biogrid_Msig_Slopes.rds")
msig_frame_string = readRDS("String_Msig_Slopes.rds")
msig_frame_string_filt = readRDS("String_Filt_Msig_Slopes.rds")

#Calculating normalized coherence
norm_coh_biogrid = calc_norm_coherence(msig_frame_biogrid, dis_slopes_biogrid_count, num_msig = num_msig, num_rand = num_rand)
norm_coh_string = calc_norm_coherence(msig_frame_string, dis_slopes_string_count, num_msig = num_msig, num_rand = num_rand)
norm_coh_string_filt = calc_norm_coherence(msig_frame_string_filt, dis_slopes_string_filt_count,num_msig = num_msig, num_rand = num_rand)

#Saving to .rds files
saveRDS(norm_coh_biogrid, "./data/Coherence_Results/coherence_biogrid.rds")
saveRDS(norm_coh_string, "./data/Coherence_Results/coherence_string.rds")
saveRDS(norm_coh_string_filt, "./data/Coherence_Results/coherence_string_filt.rds")
```

# Write general function for saving runs to .csv

```{r}
save_norm_coh = function(coherence_tab, num_msig, num_rand, run, database) {
  require(dplyr)
  fileNameOut <- file.path("./data/Coherence_Results/", paste0("coherence_msig_",database,"_",num_msig, "_rand_", num_rand, "_", run, ".csv"))
  write_csv(coherence_tab, fileNameOut)
}

```

# Save the runs in .csv form

```{r}
#Saving biogrid results to csv
save_norm_coh(norm_coh_biogrid, num_msig = num_msig, num_rand = num_rand, run = run, database = "Biogrid")

#Saving string results to csv
save_norm_coh(norm_coh_string, num_msig = num_msig, num_rand = num_rand, run = run, database = "String")

#Saving string filtered results to csv
save_norm_coh(norm_coh_string_filt, num_msig = num_msig, num_rand = num_rand, run = run, database = "String_Filtered")
```