### 2019-09-14 Analysis/Normalized_Coherence.Rmd
??? to fix

---
title: "Msigdf Biogrid" [??? Is it the right title?]
author: "Kellen Cresswell"
date: "July 19, 2019" [??? Is it the right date?]
output: pdf_document
---

# Read in functions

```{r}  
[??? What those functions do?]
source("./functions/disease2frame.R") 
source("./functions/GeneSample.R")
source("./functions/get_msig_slopes.R")

library(readr)   # install.packages("readr")
library(dplyr)   # install.packages("dplyr")
library(biomaRt) # BiocManager::install("biomaRt")
library(msigdbr) # BiocManager::install("msigdbr")
library(igraph)  # install.packages("igraph")
library(tidyr)   # install.packages("tidyr")
library(broom)   # install.packages("broom")
```

# Setting options

```{r}
# Option for number of msig pathways to take from each category
num_msig = 10
# Option for number of random networks to generate (Will generate num_rand-1 networks)
num_rand = 1001
# Construct file name to save the results
run <- 1 # Placeholder for the run ID
fileNameOut <- file.path("./data/Coherence_Results/", paste0("coherence_msig_", num_msig, "_rand_", num_rand, "_", run, ".csv"))
```

# Read in Biogrid data

```{r}
# Reading in Biogrid data
links = read_tsv("./data/BIOGRID-ALL-3.5.174.mitab.txt.gz")
links = links %>% filter(`Taxid Interactor A` == "taxid:9606" & `Taxid Interactor B` == "taxid:9606")

#Split based on Biogrid ID system then pull out gene names
protein1 = lapply(strsplit(links$`Alt IDs Interactor A`, "locuslink:"), function(x) strsplit(x[2], "\\|")[[1]][1])
protein2 = lapply(strsplit(links$`Alt IDs Interactor B`, "locuslink:"), function(x) strsplit(x[2], "\\|")[[1]][1])
protein1 = unlist(protein1)
protein2 = unlist(protein2)

#Place into matrix in form that disease2frame works on
links = links %>% mutate(protein1 =protein1, protein2 = protein2) %>% dplyr::select(protein1, protein2)
links = links %>% dplyr::select(hgnc_symbol_a = protein1,hgnc_symbol_b = protein2)
```

# Read in String data and create the String filtered dataset

```{r}
# Reading in Biogrid data
string = read_table2("./data/9606.protein.links.v11.0.txt.gz")
string = string %>% mutate(protein1 = gsub("9606.", "", protein1), protein2 = gsub("9606.", "", protein2))

# Read in ensemble protein ID, convert them to gene symbols
ensembl = useEnsembl(biomart="ensembl", dataset="hsapiens_gene_ensembl", mirror = "www")
genes = biomaRt::getBM(attributes=c('ensembl_peptide_id','hgnc_symbol'), mart = ensembl)

# Join gene names with string
string = left_join(string, genes, by = c("protein1" = "ensembl_peptide_id", "protein2" = "ensembl_peptide_id"))
string = left_join(string, genes, by = c("protein2" = "ensembl_peptide_id"))
string = string %>% dplyr::rename(hgnc_symbol_a = hgnc_symbol.x)
string = string %>% dplyr::rename(hgnc_symbol_b = hgnc_symbol.y)
string_filt = string %>% filter(combined_score>=500)
```

[??? Why do you read all data at once?]

```{r}
[??? What this function does? Write roxygen help]

get_msig_slopes = function(sub_cats, database) {
  require(msigdbr)
  msig = msigdbr()

  # Subset to include msig
  msig = msig %>% filter(gs_subcat %in% sub_cats)

  # Create empty list
  mod_over = list()
  # Set index for list entries
  index = 1
  # Loop through each pathway
  for (j in unique(msig$gs_name)) {
    #Subset to only include current pathway
    curr_msig = msig %>% filter(gs_name == j)
    #Write to table in format for disease2frame
    write.table(curr_msig$gene_symbol, file = "temp_rea_bio.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
    #Get internal/external connectivity
    diseases_curr = diseases2frame("temp_rea_bio.txt", links = links)
    #Get disease name
    diseases_curr = diseases_curr %>% mutate(Disease = j)
    mod_over[[index]] = diseases_curr
    index =  index+1
  }

  # Bind together list
  mod_over = bind_rows(mod_over)

  # Applying categories
  msig_cats  = msig %>% 
    dplyr::select(Disease = gs_name, Category = gs_subcat) %>% 
    distinct() #Get categories in order

  # Joining categories
  mod_over = left_join(mod_over, msig_cats)

  #Getting gene counts
  mod_over = mod_over %>% group_by(Disease) %>% mutate(Num_Genes = n())
  return(mod_over)
}
```

[??? What the following chunk does?]

```{r}
# Setting the three categories we want
subcats = c("CC", "CP:KEGG", "CP:REACTOME")

# Getting biogrid msig slopes [??? Do not understand, elaborate in text outside the chunk]
mod_over_biogrid = get_msig_slopes(subcats = subcats, database = links) [??? Could you refactor "links" variable]

# Getting string msig slopes
mod_over_string= get_msig_slopes(subcats = subcats, database = string)

# Getting string filtered msig slopes
mod_over_string_filt = get_msig_slopes(subcats = subcats, database = string_filt)

# Saving the files
saveRDS(mod_over_biogrid, "Biogrid_Msig_Slopes.rds")
saveRDS(mod_over_string, "String_Msig_Slopes.rds")
saveRDS(mod_over_string_filt, "String_Filt_Msig_Slopes.rds")
```

# Define function for calculating slopes and counts from internal/external edges

```{r}
[??? What this function does? Write roxygen help]
slopes_from_edges = function(edges) {
  [??? What is "KEGG" category?]
  edges %>% filter(Internal != 0) %>% filter(Category != "Random") %>% filter(Category != "KEGG") %>%
  group_by(Disease, Category) %>% mutate(Count = n()) %>% group_by(Disease, Category, Count) %>%
  do({
    mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = .)
    data.frame(Slope= coef(mod)[1])
  }) %>% arrange(Slope) %>% filter(Count>4)
  [??? Do not understand, elaborate in text outside the chunk]
}
```

# Get edges for diseases from Biogrid (Connectivity_Comparison.Rmd)

```{r}
# Biogrid
disease_overall_biogrid     <- readRDS("biogrid_edges_new.rds")
# Get Biogrid slopes and edge counts
dis_slopes_biogrid_count = slopes_from_edges(disease_overall_biogrid)

# String
disease_overall_string     <- readRDS("string_edges_new.rds")
# Get string slopes and edge counts
dis_slopes_string_count = slopes_from_edges(disease_overall_string)

# String Filtered
disease_overall_string_filt     <- readRDS("string_filt_edges_new.rds")
# Get string filtered slopes and edge counts
dis_slopes_string_filt_count = slopes_from_edges(disease_overall_string_filt)
```

# Define function for calculating normalized coherence and counts ignoring zero internal edges

```{r}
[??? What this function does? Write roxygen help]
[??? Fix indentations]
calc_norm_coherence = function(msig_frame, dis_frame, num_msig = 10, num_rand = 10) {
  Count_Sum = msig_frame %>% dplyr::filter(Internal>0) %>% 
  group_by(Disease) %>%
  mutate(Num_Genes = n()) %>% dplyr::select(Disease,Category, Num_Genes) %>% distinct() [??? What is that?]
# Splitting into list by categories [??? Fix indentation]
Count_Sum = split(Count_Sum, Count_Sum$Category)
coherence_tab = bind_rows()
[??? What this loop does?]
for (i in unique(dis_frame$Disease)) {
  # Get disease specific genes
  dis = dis_frame %>% filter(Disease == i)
  # Get 30 closest msigdf networks taking 10 from each database and bind
  close_nets = bind_rows(lapply(Count_Sum,  function(x) x %>% mutate(Diff = abs(Num_Genes-dis$Count)) %>%
    arrange(Diff) %>% head(.,num_msig))) [??? Where is 30? Should be a setting to tweak. Describe better what 30 means]
  # Subsetting msig to only include close pathways
  msig_sub = msig_frame %>% filter(Disease %in% close_nets$Disease)
  # Taking slopes for each category
  mod_msig = msig_sub %>% group_by(Disease) %>% do(tidy(
    coef(lm(I(sqrt(External)) ~ I(sqrt(Internal))-1, data = .))[[1]]
  ))
  # Take median
  mod_slope = median(mod_msig$x)
  # Creating num_rand random networks
  j = 1
  slope_list = list()
    while(j < num_rand) {
    # Empty data frame for disease networks
      diseases = bind_rows()
      # Generate random networks until there as many internal edges as current diseases
      while(nrow(diseases)<dis$Count) {
    # Generate random diseases network [??? Fix indentation]
    diseases_curr = GeneSample(dis$Count, links, directed = FALSE) [??? What this function does?]
    # Save internal and external edges
    diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = diseases_curr$internal_connectivity, External = diseases_curr$external_connectivity)
    # Filter out zeros
    diseases_curr = diseases_curr %>% filter(Internal !=0) [??? What happens if, after filtering zeros, the network will be removed? Detail]
    # Bind together non-zero internal/external edges
    diseases = bind_rows(diseases, diseases_curr)
      }
      # Save slope of current network
      slope_list[[j]] = coef(lm(I(sqrt(External)) ~ I(sqrt(Internal))-1, data = diseases))[[1]]
      # Iterate
      j = j+1

    }

      # Get median of all slopes
      rand_slope = median(unlist(slope_list))

      # Calculate coherence
      coherence = (dis$Slope - rand_slope)/(mod_slope-rand_slope)

      # Place into data frame with disease, coherence, msig normalization value and random slope normalization value
      coherence = data.frame(Disease = dis$Disease, Coherence = coherence, 
                             Msig_Norm = mod_slope, Rand_Norm = rand_slope, Count = dis$Count)
      coherence_tab = bind_rows(coherence, coherence_tab)
      return(coherence_tab)
}
}
```

# Calculating normalized coherence

```{r}
# Get read in msig pathway genes for KEGG, Reactome and GOCC
[??? Why read them if they are created above? Rename variables above. It looks like you do not need to save/read those objects at all]
msig_frame_biogrid = readRDS("Biogrid_Msig_Slopes.rds")
msig_frame_string = readRDS("String_Msig_Slopes.rds")
msig_frame_string_filt = readRDS("String_Filt_Msig_Slopes.rds")

# Calculating normalized coherence 
norm_coh_biogrid = calc_norm_coherence(msig_frame_biogrid, dis_slopes_biogrid_count, num_msig = num_msig, num_rand = num_rand)
norm_coh_string = calc_norm_coherence(msig_frame_string, dis_slopes_string_count, num_msig = num_msig, num_rand = num_rand)
norm_coh_string_filt = calc_norm_coherence(msig_frame_string_filt, dis_slopes_string_filt_count,num_msig = num_msig, num_rand = num_rand)

#Saving to .rds files
saveRDS(norm_coh_biogrid, "./data/Coherence_Results/coherence_biogrid.rds")
saveRDS(norm_coh_string, "./data/Coherence_Results/coherence_string.rds")
saveRDS(norm_coh_string_filt, "./data/Coherence_Results/coherence_string_filt.rds")
```

# Write general function for saving runs to .csv

```{r}
save_norm_coh = function(coherence_tab, num_msig, num_rand, run, database) {
  require(dplyr)
  fileNameOut <- file.path("./data/Coherence_Results/", paste0("coherence_msig_",database,"_",num_msig, "_rand_", num_rand, "_", run, ".csv"))
  write_csv(coherence_tab, fileNameOut)
}
```

# Save the runs in .csv form

```{r}
#Saving biogrid results to csv
save_norm_coh(norm_coh_biogrid, num_msig = num_msig, num_rand = num_rand, run = run, database = "Biogrid")

#Saving string results to csv
save_norm_coh(norm_coh_string, num_msig = num_msig, num_rand = num_rand, run = run, database = "String")

#Saving string filtered results to csv
save_norm_coh(norm_coh_string_filt, num_msig = num_msig, num_rand = num_rand, run = run, database = "String_Filtered")
```