### 2019-09-18 Analysis/Analysis_Tables_New.Rmd
Major rewrite to use new data

---
title: "Secondary analysis of tables"
# author: "Author's Name"
# date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: no
  html_document:
    theme: cerulean
    toc: yes
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
# Set up the environment
library(knitr)
opts_chunk$set(cache.path='cache/', fig.path='img/', cache=F, tidy=T, fig.keep='high', echo=F, dpi=100, warnings=F, message=F, comment=NA, warning=F, results='as.is', fig.width = 10, fig.height = 6) #out.width=700, 
library(pander)
panderOptions('table.split.table', Inf)
set.seed(1)
library(dplyr)
options(stringsAsFactors = FALSE)
```

```{r}
library(readr)
library(ggplot2)
library(ggrepel)
library(gridExtra)
source("https://raw.githubusercontent.com/mdozmorov/MDmisc/master/R/venn.R")
library(corrplot)
col3 <- colorRampPalette(c("blue", "white", "red"))
library(openxlsx)
library(cowplot)
```

```{r}
max_genes <- 200 # Maximum number of genes for a network to be included
min_genes <- 10  # Minimum number of genes for a network to be included
fileNameIn0  <- "../manuscript/Tables/supplementary_table_S3.csv"   # diseases and traits, coherence new KEGG, REACTOME, CC. 2019-07-14, Kellen
# iconv -f UTF-8 -t ISO-8859-1 -c supplementary_table_S3.csv > tmp.csv
mtx0 <- read_csv(fileNameIn0) %>% as.data.frame()
rownames(mtx0) <- mtx0$Disease
```

### Coherence distribution

```{r}
mtx0[, "String Filt Coherence"] %>% unlist %>% c %>% unname %>% hist
mtx0[, "String Filt Coherence"] %>% unlist %>% c %>% unname %>% summary
```

### FigS1_corr_coherence.png

**Supplementary Figure S1.[FILE: Figures/FigS1_corr_coherence.png] Normalized coherence is unaffected by network size.** Red/blue gradient and numbers represent Pearson correlation coefficients between corresponding entries.

```{r}
# Selected data for correlation
# Remove small/large networks
mtx0_cor <- mtx0[ mtx0 %>% dplyr::select(starts_with("Size")) %>% apply(., 1, min) > min_genes &
                  mtx0 %>% dplyr::select(starts_with("Size")) %>% apply(., 1, max) < max_genes, ]
# Remove size columns
mtx0_cor <- mtx0_cor %>% dplyr::select(contains("Coherence"), contains("Number"))
# Actual correlation
diff_cors <- cor(mtx0_cor, use = "pairwise.complete.obs")

# corrgram::corrgram(diff_cors, order = FALSE, upper.panel=panel.cor)
corrplot.mixed(diff_cors, lower = "number", lower.col = col3(100), upper = "circle", upper.col = col3(100), tl.cex = 1, tl.pos = "lt", tl.col = "black")
# Selected correlations
png(filename = "../manuscript/Figures/FigS1_corr_coherence.png", height = 2000, width = 2000, res = 300)
# corrgram::corrgram(diff_cors, order = FALSE, upper.panel=panel.cor, cex.labels = 4)
corrplot.mixed(diff_cors, lower = "number", lower.col = col3(100), upper = "circle", upper.col = col3(100), tl.cex = 1, tl.pos = "lt", tl.col = "black")
dev.off()
```

\pagebreak

# Summary plots

```{r}
mtx_traits_selected_dotplot <- function(category = "Intelligence", mtx_traits = mtx_traits, databases = c("Biogrid", "STRING full", "STRING filtered")) {

  mtx_traits_selected <- mtx_traits[mtx_traits$Category == category, grepl(databases, colnames(mtx_traits))]

  mtx_traits_selected_to_plot <- data.frame(Disease = rownames(mtx_traits_selected), 
                                            Pvalue = mtx_traits_selected[, grepl(paste0(databases, " P-value"), colnames(mtx_traits_selected)) ], 
                                            Coherence = mtx_traits_selected[, grepl(paste0(databases, " Coherence"), colnames(mtx_traits_selected)) ])
  # Remove noninformative rows
  mtx_traits_selected_to_plot <- mtx_traits_selected_to_plot[!is.na(mtx_traits_selected_to_plot$Coherence), ]
  # Sort by decreasing level of coherence
  mtx_traits_selected_to_plot$Disease <- factor(mtx_traits_selected_to_plot$Disease, levels = mtx_traits_selected_to_plot$Disease[order(mtx_traits_selected_to_plot$Coherence)])

  p <- ggplot(data=mtx_traits_selected_to_plot, aes(x=Disease, y=Coherence, colour = "blue", size = Coherence)) + # label = Pvalue, 
    geom_point() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    ylab("Normalized coherence") +
    # geom_text_repel(size = 4, colour = "black") +
    coord_flip() +
    # theme_cowplot(12) +
    theme(legend.position="none") +
    theme(axis.text = element_text(size = 12)) + 
    ggtitle(category)
}
```

## Traits

### Fig3_summary_traits.png

```{r results='hide'}
# STRING filtered plots only
p_intelligence <- plot(mtx_traits_selected_dotplot(category = "Intelligence", mtx_traits = mtx0, databases = "String Filt"))
p_anthropometric <- plot(mtx_traits_selected_dotplot(category = "Anthropometric trait", mtx_traits = mtx0, databases = "String Filt"))
p_cardiovascular <- plot(mtx_traits_selected_dotplot(category = "Cardiovascular trait", mtx_traits = mtx0, databases = "String Filt"))

p_eye <- plot(mtx_traits_selected_dotplot(category = "Eye trait", mtx_traits = mtx0, databases = "String Filt"))
p_metabolic <- plot(mtx_traits_selected_dotplot(category = "Metabolic trait", mtx_traits = mtx0, databases = "String Filt"))

png(filename = "../manuscript/Figures/Fig3_summary_traits.png", height = 2500, width = 5000, res = 300)
lay1 <- rbind(c(1, 2, 3))
lay2 <- rbind(c(1, 2, 3))
g1 <- arrangeGrob(grobs = list(p_anthropometric, p_cardiovascular, p_metabolic), layout_matrix = lay1)
g2 <- arrangeGrob(grobs = list(p_eye, p_intelligence), layout_matrix = lay2)
grid.arrange(g1, g2, ncol = 1)
dev.off()
```

### FigS2_summary_traits_Biogrid.png

```{r }
p_intelligence <- plot(mtx_traits_selected_dotplot(category = "Intelligence", mtx_traits = mtx0, databases = "Biogrid"))
p_anthropometric <- plot(mtx_traits_selected_dotplot(category = "Anthropometric trait", mtx_traits = mtx0, databases = "Biogrid"))
p_cardiovascular <- plot(mtx_traits_selected_dotplot(category = "Cardiovascular trait", mtx_traits = mtx0, databases = "Biogrid"))
p_eye <- plot(mtx_traits_selected_dotplot(category = "Eye trait", mtx_traits = mtx0, databases = "Biogrid"))
p_metabolic <- plot(mtx_traits_selected_dotplot(category = "Metabolic trait", mtx_traits = mtx0, databases = "Biogrid"))

png(filename = "../manuscript/Figures/FigS2_summary_traits_Biogrid.png", height = 2500, width = 6000, res = 300)
lay1 <- rbind(c(1, 2, 3))
lay2 <- rbind(c(1, 2, 3))
g1 <- arrangeGrob(grobs = list(p_anthropometric, p_cardiovascular, p_metabolic), layout_matrix = lay1)
g2 <- arrangeGrob(grobs = list(p_eye, p_intelligence), layout_matrix = lay2)
grid.arrange(g1, g2, ncol = 1)
dev.off()
```

### FigS4_summary_traits_STRING.png

```{r }
p_intelligence <- plot(mtx_traits_selected_dotplot(category = "Intelligence", mtx_traits = mtx0, databases = "String"))
p_anthropometric <- plot(mtx_traits_selected_dotplot(category = "Anthropometric trait", mtx_traits = mtx0, databases = "String"))
p_cardiovascular <- plot(mtx_traits_selected_dotplot(category = "Cardiovascular trait", mtx_traits = mtx0, databases = "String"))
p_eye <- plot(mtx_traits_selected_dotplot(category = "Eye trait", mtx_traits = mtx0, databases = "String"))
p_metabolic <- plot(mtx_traits_selected_dotplot(category = "Metabolic trait", mtx_traits = mtx0, databases = "String"))

png(filename = "../manuscript/Figures/FigS4_summary_traits_STRING.png", height = 2500, width = 6000, res = 300)
lay1 <- rbind(c(1, 2, 3))
lay2 <- rbind(c(1, 2, 3))
g1 <- arrangeGrob(grobs = list(p_anthropometric, p_cardiovascular, p_metabolic), layout_matrix = lay1)
g2 <- arrangeGrob(grobs = list(p_eye, p_intelligence), layout_matrix = lay2)
grid.arrange(g1, g2, ncol = 1)
dev.off()
```

## Diseases

### Fig4_summary_diseases.png

```{r results='hide'}
# STRING filtered plots only
p_intelligence <- plot(mtx_traits_selected_dotplot(category = "Autoimmune", mtx_traits = mtx0, databases = "String Filt"))
p_anthropometric <- plot(mtx_traits_selected_dotplot(category = "Cancer", mtx_traits = mtx0, databases = "String Filt"))
p_cardiovascular <- plot(mtx_traits_selected_dotplot(category = "Cardiovascular disease", mtx_traits = mtx0, databases = "String Filt"))
p_eye <- plot(mtx_traits_selected_dotplot(category = "Eye disease", mtx_traits = mtx0, databases = "String Filt"))
p_metabolic <- plot(mtx_traits_selected_dotplot(category = "Metabolic disease", mtx_traits = mtx0, databases = "String Filt"))
p_neurologic <- plot(mtx_traits_selected_dotplot(category = "Neurologic", mtx_traits = mtx0, databases = "String Filt"))
p_psychiatric <- plot(mtx_traits_selected_dotplot(category = "Psychiatric", mtx_traits = mtx0, databases = "String Filt"))

png(filename = "../manuscript/Figures/Fig4_summary_diseases.png", height = 2500, width = 5000, res = 300)
lay1 <- rbind(c(1, 2, 3, 4))
lay2 <- rbind(c(1, 2, 3, 4))
g1 <- arrangeGrob(grobs = list(p_intelligence, p_anthropometric, p_neurologic, p_psychiatric), layout_matrix = lay1)
g2 <- arrangeGrob(grobs = list(p_cardiovascular, p_metabolic, p_eye), layout_matrix = lay2)
grid.arrange(g1, g2, ncol = 1)
dev.off()
```

### FigS4_summary_diseases_Biogrid.png

```{r }
p_intelligence <- plot(mtx_traits_selected_dotplot(category = "Autoimmune", mtx_traits = mtx0, databases = "Biogrid"))
p_anthropometric <- plot(mtx_traits_selected_dotplot(category = "Cancer", mtx_traits = mtx0, databases = "Biogrid"))
p_cardiovascular <- plot(mtx_traits_selected_dotplot(category = "Cardiovascular disease", mtx_traits = mtx0, databases = "Biogrid"))
p_eye <- plot(mtx_traits_selected_dotplot(category = "Eye disease", mtx_traits = mtx0, databases = "Biogrid"))
p_metabolic <- plot(mtx_traits_selected_dotplot(category = "Metabolic disease", mtx_traits = mtx0, databases = "Biogrid"))
p_neurologic <- plot(mtx_traits_selected_dotplot(category = "Neurologic", mtx_traits = mtx0, databases = "Biogrid"))
p_psychiatric <- plot(mtx_traits_selected_dotplot(category = "Psychiatric", mtx_traits = mtx0, databases = "Biogrid"))

png(filename = "../manuscript/Figures/FigS4_summary_diseases_Biogrid.png", height = 3000, width = 8000, res = 300)
lay1 <- rbind(c(1, 2, 3, 4))
lay2 <- rbind(c(1, 2, 3, 4))
g1 <- arrangeGrob(grobs = list(p_intelligence, p_anthropometric, p_neurologic, p_psychiatric), layout_matrix = lay1)
g2 <- arrangeGrob(grobs = list(p_cardiovascular, p_metabolic, p_eye), layout_matrix = lay2)
grid.arrange(g1, g2, ncol = 1)
dev.off()
```

### FigS5_summary_diseases_STRING.png

```{r }
p_intelligence <- plot(mtx_traits_selected_dotplot(category = "Autoimmune", mtx_traits = mtx0, databases = "String"))
p_anthropometric <- plot(mtx_traits_selected_dotplot(category = "Cancer", mtx_traits = mtx0, databases = "String"))
p_cardiovascular <- plot(mtx_traits_selected_dotplot(category = "Cardiovascular disease", mtx_traits = mtx0, databases = "String"))
p_eye <- plot(mtx_traits_selected_dotplot(category = "Eye disease", mtx_traits = mtx0, databases = "String"))
p_metabolic <- plot(mtx_traits_selected_dotplot(category = "Metabolic disease", mtx_traits = mtx0, databases = "String"))
p_neurologic <- plot(mtx_traits_selected_dotplot(category = "Neurologic", mtx_traits = mtx0, databases = "String"))
p_psychiatric <- plot(mtx_traits_selected_dotplot(category = "Psychiatric", mtx_traits = mtx0, databases = "String"))

png(filename = "../manuscript/Figures/FigS5_summary_diseases_STRING.png", height = 3000, width = 8000, res = 300)
lay1 <- rbind(c(1, 2, 3, 4))
lay2 <- rbind(c(1, 2, 3, 4))
g1 <- arrangeGrob(grobs = list(p_intelligence, p_anthropometric, p_neurologic, p_psychiatric), layout_matrix = lay1)
g2 <- arrangeGrob(grobs = list(p_cardiovascular, p_metabolic, p_eye), layout_matrix = lay2)
grid.arrange(g1, g2, ncol = 1)
dev.off()
```

# Modularity investigation

```{r eval=FALSE}
fileNameIn0  <- "tables/Mod_Table.xlsx"   # diseases and traits, coherence new KEGG, REACTOME, CC. 2019-07-14, Kellen
# iconv -f UTF-8 -t ISO-8859-1 -c Mod_Table.csv >tmp.txt
fileNameIn0  <- "tables/tmp.txt"
mtx0 <- read_csv(fileNameIn0) %>% as.data.frame()
rownames(mtx0) <- mtx0$Disease

# Selected data for correlation
# "Disease", "Category", "Number of SNPs", "Number of genes", "Biogrid Coherence", "String Coherence", "String Filt Coherence", "Biogrid Coherence Reactome", "String Coherence Reactome", "String Filt Coherence Reactome""Biogrid Coherence CC", "String Coherence CC", "String Filt Coherence CC", "Biogrid P-value", "String P-value", "String Filt P-value" 
mtx0_cor <- mtx0 %>% dplyr::select(c("Number of SNPs", "Number of genes", "Biogrid Coherence", "String Coherence", "String Filt Coherence", "Modularity Biogrid", "Modularity String", "Modularity String Filt"))
# Remove large networks
mtx0_cor <- mtx0_cor[mtx0_cor$`Number of genes` < max_genes, ]
# Go through each pairs of columns and calculate pairwise complete observations
# pairwise_columns <- combn(ncol(mtx0_cor), 2)
# for(j in 1:ncol(pairwise_columns)) {
#   print(paste0(colnames(mtx0_cor)[pairwise_columns[1, j]], " - ", colnames(mtx0_cor)[pairwise_columns[2, j]], " : ", nrow(mtx0_cor[complete.cases(mtx0_cor[, c(pairwise_columns[1, j], pairwise_columns[2, j])]), c(pairwise_columns[1, j], pairwise_columns[2, j])])))
# }
# Actual correlation
diff_cors <- cor(mtx0_cor, use = "pairwise.complete.obs")

# corrgram::corrgram(diff_cors, order = FALSE, upper.panel=panel.cor)
corrplot.mixed(diff_cors, lower = "number", lower.col = col3(100), upper = "circle", upper.col = col3(100), tl.cex = 1, tl.pos = "lt", tl.col = "black")
# Selected correlations
png(filename = "Figures/FigS1_corr_coherence.png", height = 2000, width = 2000, res = 300)
# corrgram::corrgram(diff_cors, order = FALSE, upper.panel=panel.cor, cex.labels = 4)
corrplot.mixed(diff_cors, lower = "number", lower.col = col3(100), upper = "circle", upper.col = col3(100), tl.cex = 1, tl.pos = "lt", tl.col = "black")
dev.off()
```