### 2019-10-06 manuscript/Figures.Rmd
Make PPT of all figures

---
title: "Coherence of Molecular Mechanisms In Major Human Disease and Traits"
output: powerpoint_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

##

<center>
![**Figure 1. Degree distributions are affected by network size and configuration.** Degree distributions of A) ring networks and B) fully connected networks containing 10 and 20 nodes. While degree distributions of similarly coherent ring networks can be directly compared, degree distributions of fully connected networks depend on network size.](Figures/Fig1_network_distributions.png)
</center>

##

<center>
![**Figure 2. Network size is inversely associated with coherence.** Slopes of internal vs. external degree distributions (aka coherence) for KEGG and random networks are plotted against network size.](Figures/Fig2_Size_Slope_Plot.png)
</center>

##

<center>
![**Figure 3. Coherence estimates of traits.** Size of dots represents the level of normalized coherence (X-axis) for individual traits (Y-axis).](Figures/Fig3_summary_traits.png)
</center>

##

<center>
![**Figure 4. Coherence estimates of diseases.** Size of dots represents the level of normalized coherence (X-axis) for individual diseases (Y-axis).](Figures/Fig4_summary_diseases.png)
</center>

##

<center>
![**Supplementary Figure S1. Normalization of coherence estimates alleviates its network size dependence.** Red/blue gradient and numbers represent Pearson correlation coefficients for each pairwise comparison of the number of SNPs, genes, and the normalized coherence estimates using the corresponding PPI databases.](Figures/FigS1_corr_coherence.png)
</center>

##

<center>
![**Supplementary Figure S2. Coherence estimates of traits using Biogrid as a reference PPI database.** Size of dots represents the level of normalized coherence (X-axis) for individual traits (Y-axis). Plots are faceted by categories. Missing entries indicate that, for a given trait, a network could not be built and the coherence cannot be estimated.](Figures/FigS2_summary_traits_Biogrid.png)
</center>

##

<center>
![**Supplementary Figure S3. Coherence estimates of traits using STRING as a reference of PPI database.** See legend for Supplementary Figure S2.](Figures/FigS3_summary_traits_STRING.png)
</center>

##

<center>
![**Supplementary Figure S4. Coherence estimates of diseases using Biogrid as a reference PPI database.** See legend for Supplementary Figure S2.](Figures/FigS4_summary_diseases_Biogrid.png)
</center>

##

<center>
![**Supplementary Figure S5. Coherence estimates of diseases using STRING as a reference PPI database.** See legend for Supplementary Figure S2.](Figures/FigS5_summary_diseases_STRING.png)
</center>