### 2019-10-16 Analysis/07_Modularity_Analysis.Rmd
add modularity

---
title: "07_Modularity_Analysis"
author: "Kellen Cresswell"
date: "October 15, 2019"
output: pdf_document
---

```{r}
require(dplyr)
options(stringsAsFactors = FALSE)
```

```{r}
#Read in modularity

biogrid_mod = as.data.frame(readRDS("./data/Modularity_Results/biogrid_mod.rds"))
string_mod = as.data.frame(readRDS("./data/Modularity_Results/string_mod.rds"))
string_filt_mod = as.data.frame(readRDS("./data/Modularity_Results/string_filt_mod.rds"))
```

```{r}
#Read in coherence table

coherence_table = read.csv(".././manuscript/Tables/supplementary_table_S3.csv")

#Putting together frame for merging

mod_frame_b_s = full_join(biogrid_mod, string_mod, "i")

mod_frame = full_join(mod_frame_b_s, string_filt_mod, "i")

#Fixing column names

mod_frame = mod_frame %>% dplyr::rename(Disease = i, Biogrid.Modularity = mod.x, String.Modularity=mod.y, String.Filt.Modularity = mod)

#Join to coherence_table

coherence_table = left_join(coherence_table, mod_frame, "Disease")

#Fix mod cutoff being 5 and p-value being 10 for filtered

coherence_table = coherence_table %>% mutate(String.Filt.Modularity = ifelse(is.na(String.Filt.P.value), NA, String.Filt.Modularity   ))

#Remove . added by read.csv

colnames(coherence_table) = gsub("\\.", " ",colnames(coherence_table))
```

```{r}
#Write to table
write.csv(coherence_table, ".././manuscript/Tables/supplementary_table_S3.csv", 
          row.names = FALSE)
```