### 2020-02-25 Analysis/08_Generate_Internal_Distributions.Rmd
Add degree comparison analysis

---
title: "Density_Plots"
author: "Kellen Cresswell"
date: "February 22, 2020"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r}
require(dplyr)
require(ggplot2)
library(readr)
source("functions/GeneSample.R")
library(cowplot)
library(gridExtra)
```

```{r}
#Repeating for string

data_dir = "./data/"

#Read in new string database

string = read_table2(paste0(data_dir,"9606.protein.links.v11.0.txt.gz"))

string = string %>% mutate(protein1 = gsub("9606.", "", protein1), 
                             protein2 = gsub("9606.", "", protein2))

#Read in ensemble protein ID

ensembl = biomaRt::useEnsembl(biomart="ensembl", dataset="hsapiens_gene_ensembl", mirror = "www")

#Get gene set
genes = biomaRt::getBM(attributes=c('ensembl_peptide_id','hgnc_symbol'), mart = ensembl)

#Join gene names with string

string = left_join(string, genes, by = c("protein1" = "ensembl_peptide_id", "protein2" = "ensembl_peptide_id"))

#Rename to work with our functions

string = string %>% dplyr::rename(hgnc_symbol_a = hgnc_symbol)

#Repeat for second set of proteins

string = left_join(string, genes, by = c("protein2" = "ensembl_peptide_id"))

string = string %>% dplyr::rename(hgnc_symbol_b = hgnc_symbol)

string_filt = string %>% filter(combined_score>=500)

dim(string)
```

```{r}
disease_overall = readRDS("biogrid_edges_new.rds")
disease_overall_string = readRDS("string_edges_new.rds")
disease_overall_string_filt = readRDS("string_filt_edges_new.rds")
```

```{r}
coherence = readRDS("data/Coherence_Results/coherence_biogrid.rds") %>% dplyr::mutate(High = ifelse(Coherence>.85, "High", ifelse(Coherence<.58, "Low", "Neither"))) %>% dplyr::select(Disease, High)
coherence_string = readRDS("data/Coherence_Results/coherence_string.rds") %>% dplyr::mutate(High = ifelse(Coherence>.85, "High", ifelse(Coherence<.58, "Low", "Neither"))) %>% dplyr::select(Disease, High)
coherence_string_filt =  readRDS("data/Coherence_Results/coherence_string_filt.rds") %>% dplyr::mutate(High = ifelse(Coherence>.85, "High", ifelse(Coherence<.58, "Low", "Neither"))) %>% dplyr::select(Disease, High)
```

```{r}
counts_biogrid = disease_overall %>% group_by(Disease) %>% filter(Internal !=0) %>% summarise(Count = n())

counts_string = disease_overall_string %>% group_by(Disease) %>% filter(Internal !=0) %>% summarise(Count = n())

counts_string_filt = disease_overall_string_filt %>% group_by(Disease) %>% filter(Internal !=0) %>% summarise(Count = n())
```

```{r}
disease_overall = left_join(disease_overall, coherence) %>% dplyr::filter(High %in% c("High", "Low"))
disease_overall_string = left_join(disease_overall_string, coherence_string)  %>% dplyr::filter(High %in% c("High", "Low"))
disease_overall_string_filt = left_join(disease_overall_string_filt, coherence_string_filt)  %>% dplyr::filter(High %in% c("High", "Low"))
```

```{r}
disease_overall = left_join(disease_overall, counts_biogrid)
disease_overall_string = left_join(disease_overall_string, counts_string)
disease_overall_string_filt = left_join(disease_overall_string_filt, counts_string_filt)

```

```{r}
#Generate random networks equivalent to high coherence (STRING)
random_string_high = bind_rows()

for (i in unique((disease_overall_string %>% filter(High == "High"))$Count)) {
 diseases_curr = GeneSample(i, string, directed = FALSE)
 diseases_curr = data.frame(Genes = names(diseases_curr$internal_connectivity), Internal = (diseases_curr$internal_connectivity), External = (diseases_curr$external_connectivity), Disease = "Random", Category = "Random", High = "Random", Count = i)
 diseases_curr  = diseases_curr %>% filter(Internal != 0)
  random_string_high = bind_rows(diseases_curr, random_string_high)
}

#Generate random networks equivalent to low coherence (STRING)
random_string_low = bind_rows()

for (i in unique((disease_overall_string %>% filter(High == "Low"))$Count)) {
 diseases_curr = GeneSample(i, string, directed = FALSE)
 diseases_curr = data.frame(Genes = names(diseases_curr$internal_connectivity), Internal = (diseases_curr$internal_connectivity), External = (diseases_curr$external_connectivity), Disease = "Random", Category = "Random", High = "Random", Count = i)
 diseases_curr  = diseases_curr %>% filter(Internal != 0)
  random_string_low = bind_rows(diseases_curr, random_string_low)
}

#Generate random networks equivalent to high coherence (STRING Filtered)
random_string_filt_high = bind_rows()

for (i in unique((disease_overall_string_filt %>% filter(High == "High"))$Count)) {
 diseases_curr = GeneSample(i, string_filt, directed = FALSE)
 diseases_curr = data.frame(Genes = names(diseases_curr$internal_connectivity), Internal = (diseases_curr$internal_connectivity), External = (diseases_curr$external_connectivity), Disease = "Random", Category = "Random", High = "Random", Count = i)
 diseases_curr  = diseases_curr %>% filter(Internal != 0)
  random_string_filt_high = bind_rows(diseases_curr, random_string_filt_high)
}

#Generate random networks equivalent to low coherence (STRING Filtered)
random_string_filt_low = bind_rows()

for (i in unique((disease_overall_string_filt %>% filter(High == "Low"))$Count)) {
 diseases_curr = GeneSample(i, string_filt, directed = FALSE)
 diseases_curr = data.frame(Genes = names(diseases_curr$internal_connectivity), Internal = (diseases_curr$internal_connectivity), External = (diseases_curr$external_connectivity), Disease = "Random", Category = "Random", High = "Random", Count = i)
 diseases_curr  = diseases_curr %>% filter(Internal != 0)
  random_string_filt_low = bind_rows(diseases_curr, random_string_filt_low)
}

#Adding random to string and string filtered
disease_overall_string = bind_rows(disease_overall_string, random_string_high, random_string_low)

disease_overall_string_filt = bind_rows(disease_overall_string_filt, random_string_filt_high, random_string_filt_low)
```

```{r}
#Performing wilcox test of internal edges
string_wilcox = pairwise.wilcox.test(disease_overall_string$Internal, disease_overall_string$High, "bonferroni")
string_wilcox_filt = pairwise.wilcox.test(disease_overall_string_filt$Internal, disease_overall_string_filt$High, "bonferroni")

#Performing wilcox test of external edges
string_wilcox_ex = pairwise.wilcox.test(disease_overall_string$External, disease_overall_string$High, "bonferroni")
string_wilcox_filt_ex = pairwise.wilcox.test(disease_overall_string_filt$External, disease_overall_string_filt$High, "bonferroni")

#Performing KS test of internal edges
# source("https://raw.githubusercontent.com/happyrabbit/DataScienceR/master/R/pairwise_ks_test.R")
# string_ks = pairwise_ks_test(disease_overall_string$Internal, disease_overall_string$High)
# string_ks_filt = pairwise_ks_test(disease_overall_string_filt$Internal, disease_overall_string_filt$High)
# string_ks
# string_ks_filt

# Save in Supplementary Table S4 that will be Additional File 9
x <- list(data.frame(Coherence = c("Low", "Random"), string_wilcox$p.value), 
          data.frame(Coherence = c("Low", "Random"), string_wilcox_filt$p.value), 
          data.frame(Coherence = c("Low", "Random"), string_wilcox_ex$p.value), 
          data.frame(Coherence = c("Low", "Random"), string_wilcox_filt_ex$p.value))
names(x) <- c("STRING_internal", "STRINGfilt_internal", "STIRNG_external",  "STIRNGfilt_external")
writexl::write_xlsx(x, path = "../manuscript/Tables/supplementary_table_S4.xlsx")
```

```{r}
#Internal edges 
# gg_biogrid = ggplot(disease_overall %>% dplyr::filter(Internal != 0), aes(x = Internal, fill = High)) + geom_density(alpha = 0.1) + labs(fill = "Coherence",x = "Internal Edges")
# plot(gg_biogrid)
# ggsave(gg_biogrid + theme_bw(20), file= './figures/Internal_Distribution_Biogrid.png', units = "in", width = 3.5, height = 2.5, scale = 2.8)

gg_string = ggplot(disease_overall_string %>% dplyr::filter(Internal != 0), aes(x = Internal, fill = High)) + geom_density(alpha = 0.3) + labs(fill = "Coherence", x = "Internal Edges", y = "Density") + xlim(0, 75) + theme_cowplot() + theme(legend.position = c(0.6, 0.6)) + ggtitle("STRING") + theme(plot.title = element_text(size=12, hjust = 0.5, margin = margin(t = 10, b = -10)))
plot(gg_string)
# ggsave(gg_string, file= './figures/Internal_Distribution_String.png', units = "in", width = 4, height = 3)

gg_string_filt = ggplot(disease_overall_string_filt %>% dplyr::filter(Internal != 0), aes(x = Internal, fill = High)) + geom_density(alpha = 0.3) + labs(fill = "Coherence", x = "Internal Edges", y = "Density") + xlim(0, 25) + theme_cowplot() + theme(legend.position = "none") + ggtitle("STRING Filtered") + theme(plot.title = element_text(size=12, hjust = 0.5, margin = margin(t = 10, b = -10)))
# plot(gg_string_filt)
# ggsave(gg_string_filt, file= './figures/Internal_Distribution_String_Filt.png', units = "in", width = 4, height = 3)

#External edges
gg_string_ex = ggplot(disease_overall_string, aes(x = External, fill = High)) + geom_density(alpha = 0.3) + labs(fill = "Coherence", x = "External Edges", y = "Density") + xlim(0, 4000) + theme_cowplot() + theme(legend.position = "none") + ggtitle("STRING") + theme(plot.title = element_text(size=12, hjust = 0.5, margin = margin(t = 10, b = -10)))
# print(gg_string_ex)
# ggsave(gg_string_ex, file= './figures/External_Distribution_String.png', units = "in", width = 4, height = 3)

gg_string_filt_ex = ggplot(disease_overall_string_filt, aes(x = External, fill = High)) + geom_density(alpha = 0.3) + labs(fill = "Coherence", x = "External Edges", y = "Density") + xlim(0, 500) + theme_cowplot() + theme(legend.position = "none") + ggtitle("STRING Filtered") + theme(plot.title = element_text(size=12, hjust = 0.5, margin = margin(t = 10, b = -10)))
# print(gg_string_filt_ex)
# ggsave(gg_string_filt_ex, file= './figures/External_Distribution_String_Filt.png', units = "in", width = 4, height = 3)

png(filename = "../manuscript/Figures/Fig3_degree_distributions.png", height = 5, width = 7, units = "in", res = 300)
lay1 <- rbind(c(1, 2))
lay2 <- rbind(c(1, 2))
g1 <- arrangeGrob(grobs = list(gg_string, gg_string_filt), layout_matrix = lay1)
g2 <- arrangeGrob(grobs = list(gg_string_ex, gg_string_filt_ex), layout_matrix = lay2)
grid.arrange(g1, g2, ncol = 1)
dev.off()
```